---
title: "Q4: Trait-Based Model Prediction"
author: "Indra Boving"
date: "2026-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(brms)
library(tidyverse)
library(here)
```


2) If you were to take all the significant relationships of model parameters with traits, could you build a multivariate model that predicts LFMmin, and then compare it to observed? The value would be for modelers who have functional trait databases, to show how close we can get in predictions based only on the functional traits

#Read in data: 

```{r}
concept_df <- read_csv(here("data", "concept_df.csv")) %>% 
  filter(group == "species") %>% 
  select(mwp, ilfm, species, study) 

q3_df <- read_csv(here("data", "q3_df.csv"))

q4_df <- merge(concept_df, q3_df, by = c("species")) %>%
  filter(!is.na(sla)) %>%
  mutate(
    psi_z = as.numeric(scale(mwp)),   # or psi_md if that’s the column name
    sla_z = as.numeric(scale(sla))
  )

q4_df %>% write_csv(here("data", "q4_df.csv"))
```

#Table S5: Final model selection
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(Metrics)
library(knitr)
library(kableExtra)
library(here)

# ---------------------------------------------------------
# 1. Prepare the Master Dataset
# ---------------------------------------------------------
# We must drop NAs for all variables used in ANY of the 9 models
# so they are all compared on the exact same dataset.
df_mod <- q4_df %>%
  filter(plc50_stem >= -8 & plc50_stem <= -1) %>%
  filter(sla > 0 & sla < 100) %>% 
  filter(
    !is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), 
    !is.na(ldmc), !is.na(psi_tlp), !is.na(functional_type), !is.na(species)
  ) %>%
  mutate(
    mwp_z    = as.numeric(scale(mwp)),
    p50_z    = as.numeric(scale(plc50_stem)),
    sla_z    = as.numeric(scale(sla)),
    tlp_z    = as.numeric(scale(psi_tlp)),
    ldmc_z   = as.numeric(scale(ldmc)),
    functional_type = as.factor(functional_type)
  )

# VIF Check
library(car)
species_traits <- df_mod %>% select(species, sla, ldmc) %>% distinct()
cat("SLA-LDMC correlation:", cor(species_traits$sla, species_traits$ldmc), "\n")

mod_vif_check <- lmer(
  ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),
  data = df_mod, REML = TRUE
)
cat("VIF:\n")
print(vif(mod_vif_check))

# ---------------------------------------------------------
# 2. Define Your 9 A Priori Models
# ---------------------------------------------------------
formulas <- list(
  "P50 + SLA + LDMC + TLP" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * tlp_z + (1 | species),
  
  "P50 + SLA + LDMC + TLP + F. Type" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * tlp_z + (1 | species),
  
  "P50 + SLA + LDMC + F. Type" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "TLP + SLA + LDMC + F. Type" = ilfm ~ mwp_z * tlp_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "P50 + SLA + LDMC" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),
  
  "SLA + LDMC + F. Type" = ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "LDMC + F. Type" = ilfm ~ mwp_z * ldmc_z +  functional_type * mwp_z + (1 | species),
  
  "SLA + F. Type" = ilfm ~ mwp_z * sla_z + functional_type * mwp_z + (1 | species),
  
  "SLA Only" = ilfm ~ mwp_z * sla_z + (1 | species),
  
  # P50 + TLP + F. Type (Hydraulics only, no leaf economics)
  "TLP + P50 + F. Type" = ilfm ~ mwp_z * p50_z + mwp_z * tlp_z + functional_type * mwp_z + (1 | species),
  
  "TLP Only" = ilfm ~ mwp_z * tlp_z + (1 | species),
  
  "P50 Only" = ilfm ~ mwp_z * p50_z+ (1 | species),
  
  "LDMC Only" = ilfm ~ mwp_z * ldmc_z+ (1 | species),
  
  "F.Type Only" = ilfm ~ mwp_z * functional_type + (1 | species),
  
  "WP Only (Null)" = ilfm ~ mwp_z + (1 | species)
)

# ---------------------------------------------------------
# 3. Evaluate Models and Run Cross-Validation
# ---------------------------------------------------------
results_list <- list()
species_list <- unique(df_mod$species)

cat("Running evaluations and cross-validation for 9 hypothesis-driven models...\n")

for(model_name in names(formulas)) {
  
  current_formula <- formulas[[model_name]]
  
  # A. Fit Full Model (REML = FALSE is required to compare fixed effects with AIC)
  full_mod <- lmer(current_formula, data = df_mod, REML = FALSE)
  
  # B. Calculate AIC & In-Sample R2
  mod_aic <- AIC(full_mod)
  in_sample_preds <- predict(full_mod)
  in_sample_r2 <- cor(df_mod$ilfm, in_sample_preds)^2
  
  # C. Leave-One-Species-Out CV Loop
  cv_preds <- numeric(nrow(df_mod))
  
  for(sp in species_list) {
    train_data <- df_mod %>% filter(species != sp)
    test_data  <- df_mod %>% filter(species == sp)
    
    cv_mod <- lmer(current_formula, data = train_data, REML = FALSE)
    cv_preds[df_mod$species == sp] <- predict(cv_mod, newdata = test_data, allow.new.levels = TRUE)
  }
  cv_r2 <- cor(df_mod$ilfm, cv_preds)^2
  
  # D. Store Results
  results_list[[model_name]] <- data.frame(
    Model = model_name,
    AIC = mod_aic,
    In_Sample_R2 = in_sample_r2,
    Cross_Val_R2 = cv_r2
  )
}

# ---------------------------------------------------------
# 4. Format the Dataframe
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

min_aic <- min(model_comparison_df$AIC)

final_df <- model_comparison_df %>%
  mutate(Delta_AIC = AIC - min_aic) %>%
  arrange(Delta_AIC) %>%
  # Round numbers for clean presentation
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  # Clean up headers
  rename(
    `Candidate Model` = Model,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

rownames(final_df) <- NULL

# ---------------------------------------------------------
# 5. Create and Save the Publication kable Table
# ---------------------------------------------------------
pub_table <- kable(final_df, format = "html", escape = FALSE, align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  # Highlight the best model row in bold
  row_spec(1, bold = TRUE) 

# Print to Viewer
pub_table

# Save as an HTML file for easy copy/pasting into Word
write.csv(final_df, here("results_final", "tables", "Table_S07_model_comparison.csv"), row.names = FALSE)
tryCatch(save_kable(pub_table, file = here("results_final", "tables", "Table_S07_model_comparison.html")), error = function(e) message("Note: HTML table skipped (pandoc not available)"))

cat("Done! Table saved to figures/Table_Hypothesis_Models.html\n")

# ---------------------------------------------------------
# 4. Filter, Format, and Rank
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

min_aic <- min(model_comparison_df$AIC)

final_df <- model_comparison_df %>%
  mutate(Delta_AIC = AIC - min_aic) %>%
  # Explicitly keep models within 10 AIC, PLUS your two baselines
  filter(Delta_AIC <= 10 | Model %in% c("WP Only (Null)", "F. Type Only")) %>%
  arrange(Delta_AIC) %>%
  # Round numbers for clean presentation
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  # Clean up headers
  rename(
    `Candidate Model` = Model,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

rownames(final_df) <- NULL

# ---------------------------------------------------------
# 5. Create and Save the Publication kable Table
# ---------------------------------------------------------
pub_table <- kable(final_df, format = "html", escape = FALSE, align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  row_spec(1, bold = TRUE) 

# Print to Viewer
pub_table

# Save as an HTML file for easy copy/pasting into Word
write.csv(final_df, here("results_final", "tables", "Table_S07_model_comparison_filtered.csv"), row.names = FALSE)
tryCatch(save_kable(pub_table, file = here("results_final", "tables", "Table_S07_model_comparison_filtered.html")), error = function(e) message("Note: HTML table skipped (pandoc not available)"))

cat("Done! Table saved to figures/Table_Hypothesis_Models_Filtered.html\n")
```


####v2: Colored points 
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(Metrics) # for calculating RMSE and R-squared
library(here)

# ---------------------------------------------------------
# 1. Load, clean data, and assign plot groups
# ---------------------------------------------------------
df <- q4_df

df_mod <- df %>%
  # Drop missing values
  filter(!is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), !is.na(ldmc), !is.na(species), !is.na(psi_tlp)) %>%
  # Standardize continuous predictors (z-scores)
  mutate(
    mwp_z = as.numeric(scale(mwp)),
    p50_z = as.numeric(scale(plc50_stem)),
    sla_z = as.numeric(scale(sla)),
    tlp_z = as.numeric(scale(psi_tlp)),
    ldmc_z = as.numeric(scale(ldmc))
  ) %>%
  # Create the grouping variable for the plot colors
  mutate(
    plot_group = case_when(
      grepl("gymnosperm", functional_type, ignore.case = TRUE) ~ "Conifer",
      grepl("sal", species, ignore.case = TRUE) ~ "Salvia",
      TRUE ~ "Other Angiosperm"
    ),
    # Make it a factor to control the legend order
    plot_group = factor(plot_group, levels = c("Conifer", "Salvia", "Other Angiosperm"))
  )

# ---------------------------------------------------------
# 2. Frequentist Framework (lmer)
# ---------------------------------------------------------
freq_mod <- lmer(
  ilfm ~  mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * p50_z + mwp_z * tlp_z+ (1 | species), 
  data = df_mod, 
  REML = TRUE
)

# summary() will now give you p-values
summary(freq_mod)

# ---------------------------------------------------------
# 3. Model Predictions (In-Sample and Out-of-Sample)
# ---------------------------------------------------------
# In-Sample
df_mod$predicted_ilfm <- predict(freq_mod)
in_sample_r2 <- cor(df_mod$ilfm, df_mod$predicted_ilfm)^2
in_sample_rmse <- rmse(df_mod$ilfm, df_mod$predicted_ilfm)

# Out-of-Sample (Leave-One-Species-Out)
df_mod$cv_predicted_ilfm <- NA
species_list <- unique(df_mod$species)

for(sp in species_list) {
  train_data <- df_mod %>% filter(species != sp)
  test_data <- df_mod %>% filter(species == sp)
  
  temp_mod <- lmer(
    ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * p50_z + mwp_z * tlp_z +(1 | species), 
    data = train_data,
    REML = FALSE 
  )
  
  preds <- predict(temp_mod, newdata = test_data, allow.new.levels = TRUE)
  df_mod$cv_predicted_ilfm[df_mod$species == sp] <- preds
}

cv_r2 <- cor(df_mod$ilfm, df_mod$cv_predicted_ilfm)^2
cv_rmse <- rmse(df_mod$ilfm, df_mod$cv_predicted_ilfm)

# ---------------------------------------------------------
# 4. Build Publication-Ready Plots
# ---------------------------------------------------------
# Find the absolute min and max across ALL predictions and observations 
min_val <- min(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)
max_val <- max(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)

anno_text_insample <- sprintf("In-Sample R2 = %.2f\nRMSE = %.2f", in_sample_r2, in_sample_rmse)
anno_text_cv <- sprintf("Predictive R2 = %.2f\nRMSE = %.2f", cv_r2, cv_rmse)

# Custom color palette
my_colors <- c(
  "Conifer" = "#636B2F",       # Light Green
  "Salvia" = "#F768A1",           # Pink
  "Other Angiosperm" = "grey70"   # Default Grey
)
 df_mod <- df_mod %>%
  # Make sure the label is "Conifer" instead of "Gymnosperm"
  mutate(
    plot_group = str_replace(plot_group, "Gymnosperm", "Conifer"),
    plot_group = factor(plot_group, levels = c("Conifer", "Salvia", "Other Angiosperm"))
  ) %>%
  # Sorting by desc(plot_group) puts "Other Angiosperm" first in the dataframe
  arrange(desc(plot_group))
# Panel A: In-Sample Plot
plot_insample_pub <- ggplot(df_mod, aes(x = predicted_ilfm, y = ilfm)) +
  # Map the fill to our new plot_group
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_insample, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (In-Sample)",
    y = "Observed iLFM",
    fill = "Plant Group" # Clean legend title
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = "none" # Hide legend on the first plot so it doesn't duplicate
  )

# Panel B: Cross-Validation Plot
plot_cv_pub <- ggplot(df_mod, aes(x = cv_predicted_ilfm, y = ilfm)) +
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_cv, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (Unseen Species)",
    y = "Observed iLFM",
    fill = "Plant Group"
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) 
  )

# ---------------------------------------------------------
# 5. Combine and Save
# ---------------------------------------------------------
# Stitch together with patchwork and add A/B tags
# collect_guides() ensures we only have one clean legend for the whole figure
combined_plot <- plot_insample_pub + plot_cv_pub + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = "collect") 

combined_plot

ggsave(
  filename = here("results_final", "main", "Fig_10_observed_vs_predicted.png"), 
  plot     = combined_plot,
  width    = 13,   # Added an extra inch of width to make room for the legend
  height   = 6,   
  units    = "in",
  dpi      = 300
)
```

##Bayesian approach: 


#Version with just angiosperms: 

###Table S6: Final model selection, no gymnos:
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(Metrics)
library(knitr)
library(kableExtra)
library(here)

# ---------------------------------------------------------
# 1. Prepare the Master Dataset
# ---------------------------------------------------------
# We must drop NAs for all variables used in ANY of the 9 models
# so they are all compared on the exact same dataset.
df_mod <- q4_df %>%
  filter(plc50_stem >= -8 & plc50_stem <= -1) %>%
  filter(sla > 0 & sla < 100) %>% 
  filter(
    !is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), 
    !is.na(ldmc), !is.na(psi_tlp), !is.na(functional_type), !is.na(species)
  ) %>%
  mutate(
    mwp_z    = as.numeric(scale(mwp)),
    p50_z    = as.numeric(scale(plc50_stem)),
    sla_z    = as.numeric(scale(sla)),
    tlp_z    = as.numeric(scale(psi_tlp)),
    ldmc_z   = as.numeric(scale(ldmc)),
    functional_type = as.factor(functional_type)
  ) %>% 
  filter(functional_type == "angiosperm")

# ---------------------------------------------------------
# 2. Define Your 9 A Priori Models
# ---------------------------------------------------------
formulas <- list(
  "P50 + SLA + LDMC + TLP" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * tlp_z + (1 | species),

  "P50 + SLA + LDMC" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),

  "TLP + SLA + LDMC" = ilfm ~ mwp_z * tlp_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),

  "SLA + LDMC" = ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),

  "TLP + P50" = ilfm ~ mwp_z * p50_z + mwp_z * tlp_z + (1 | species),

  "SLA + TLP" = ilfm ~ mwp_z * sla_z + mwp_z * tlp_z + (1 | species),

  "LDMC + TLP" = ilfm ~ mwp_z * ldmc_z + mwp_z * tlp_z + (1 | species),

  "SLA + P50" = ilfm ~ mwp_z * sla_z + mwp_z * p50_z + (1 | species),

  "LDMC + P50" = ilfm ~ mwp_z * ldmc_z + mwp_z * p50_z + (1 | species),

  "SLA Only" = ilfm ~ mwp_z * sla_z + (1 | species),

  "TLP Only" = ilfm ~ mwp_z * tlp_z + (1 | species),

  "P50 Only" = ilfm ~ mwp_z * p50_z + (1 | species),

  "LDMC Only" = ilfm ~ mwp_z * ldmc_z + (1 | species),

  "WP Only (Null)" = ilfm ~ mwp_z + (1 | species)
)

# ---------------------------------------------------------
# 3. Evaluate Models and Run Cross-Validation
# ---------------------------------------------------------
results_list <- list()
species_list <- unique(df_mod$species)

cat("Running evaluations and cross-validation for 9 hypothesis-driven models...\n")

for(model_name in names(formulas)) {
  
  current_formula <- formulas[[model_name]]
  
  # A. Fit Full Model (REML = FALSE is required to compare fixed effects with AIC)
  full_mod <- lmer(current_formula, data = df_mod, REML = FALSE)
  
  # B. Calculate AIC & In-Sample R2
  mod_aic <- AIC(full_mod)
  in_sample_preds <- predict(full_mod)
  in_sample_r2 <- cor(df_mod$ilfm, in_sample_preds)^2
  
  # C. Leave-One-Species-Out CV Loop
  cv_preds <- numeric(nrow(df_mod))
  
  for(sp in species_list) {
    train_data <- df_mod %>% filter(species != sp)
    test_data  <- df_mod %>% filter(species == sp)
    
    cv_mod <- lmer(current_formula, data = train_data, REML = FALSE)
    cv_preds[df_mod$species == sp] <- predict(cv_mod, newdata = test_data, allow.new.levels = TRUE)
  }
  cv_r2 <- cor(df_mod$ilfm, cv_preds)^2
  
  # D. Store Results
  results_list[[model_name]] <- data.frame(
    Model = model_name,
    AIC = mod_aic,
    In_Sample_R2 = in_sample_r2,
    Cross_Val_R2 = cv_r2
  )
}

# ---------------------------------------------------------
# 4. Format the Dataframe
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

min_aic <- min(model_comparison_df$AIC)

final_df <- model_comparison_df %>%
  mutate(Delta_AIC = AIC - min_aic) %>%
  arrange(Delta_AIC) %>%
  # Round numbers for clean presentation
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  # Clean up headers
  rename(
    `Candidate Model` = Model,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

rownames(final_df) <- NULL

# ---------------------------------------------------------
# 5. Create and Save the Publication kable Table
# ---------------------------------------------------------
pub_table <- kable(final_df, format = "html", escape = FALSE, align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  # Highlight the best model row in bold
  row_spec(1, bold = TRUE) 

# Print to Viewer
pub_table

# Save as an HTML file for easy copy/pasting into Word
write.csv(final_df, here("results_final", "tables", "Table_S07_model_comparison_angiosperms.csv"), row.names = FALSE)
tryCatch(save_kable(pub_table, file = here("results_final", "tables", "Table_S07_model_comparison_angiosperms.html")), error = function(e) message("Note: HTML table skipped (pandoc not available)"))

cat("Done! Table saved to figures/Table_Hypothesis_Models_nogymno.html\n")
```

####Fig nogymno: Colored points 
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(Metrics) # for calculating RMSE and R-squared
library(here)

# ---------------------------------------------------------
# 1. Load, clean data, and assign plot groups
# ---------------------------------------------------------
df <- q4_df %>%
  filter(functional_type == "angiosperm")

df_mod <- df %>%
  filter(!is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), !is.na(ldmc), !is.na(species), !is.na(psi_tlp)) %>%
  mutate(
    mwp_z = as.numeric(scale(mwp)),
    p50_z = as.numeric(scale(plc50_stem)),
    sla_z = as.numeric(scale(sla)),
    tlp_z = as.numeric(scale(psi_tlp)),
    ldmc_z = as.numeric(scale(ldmc))
  ) %>%
  mutate(
    plot_group = case_when(
      grepl("sal", species, ignore.case = TRUE) ~ "Salvia",
      TRUE ~ "Other Angiosperm"
    ),
    plot_group = factor(plot_group, levels = c("Salvia", "Other Angiosperm"))
  )

# ---------------------------------------------------------
# 2. Frequentist Framework (lmer)
# ---------------------------------------------------------
freq_mod <- lmer(
  ilfm ~  mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * p50_z + mwp_z * tlp_z+ (1 | species), 
  data = df_mod, 
  REML = TRUE
)

# summary() will now give you p-values
summary(freq_mod)

# ---------------------------------------------------------
# 3. Model Predictions (In-Sample and Out-of-Sample)
# ---------------------------------------------------------
# In-Sample
df_mod$predicted_ilfm <- predict(freq_mod)
in_sample_r2 <- cor(df_mod$ilfm, df_mod$predicted_ilfm)^2
in_sample_rmse <- rmse(df_mod$ilfm, df_mod$predicted_ilfm)

# Out-of-Sample (Leave-One-Species-Out)
df_mod$cv_predicted_ilfm <- NA
species_list <- unique(df_mod$species)

for(sp in species_list) {
  train_data <- df_mod %>% filter(species != sp)
  test_data <- df_mod %>% filter(species == sp)
  
  temp_mod <- lmer(
    ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z  + mwp_z * p50_z + mwp_z * tlp_z +(1 | species), 
    data = train_data,
    REML = FALSE 
  )
  
  preds <- predict(temp_mod, newdata = test_data, allow.new.levels = TRUE)
  df_mod$cv_predicted_ilfm[df_mod$species == sp] <- preds
}

cv_r2 <- cor(df_mod$ilfm, df_mod$cv_predicted_ilfm)^2
cv_rmse <- rmse(df_mod$ilfm, df_mod$cv_predicted_ilfm)

# ---------------------------------------------------------
# 4. Build Publication-Ready Plots
# ---------------------------------------------------------
# Find the absolute min and max across ALL predictions and observations 
min_val <- min(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)
max_val <- max(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)

anno_text_insample <- sprintf("In-Sample R2 = %.2f\nRMSE = %.2f", in_sample_r2, in_sample_rmse)
anno_text_cv <- sprintf("Predictive R2 = %.2f\nRMSE = %.2f", cv_r2, cv_rmse)

# Custom color palette
my_colors <- c(
  "Conifer" = "#636B2F",       # Light Green
  "Salvia" = "#F768A1",           # Pink
  "Other Angiosperm" = "grey70"   # Default Grey
)
 df_mod <- df_mod %>%
  # Make sure the label is "Conifer" instead of "Gymnosperm"
  mutate(
    plot_group = str_replace(plot_group, "Gymnosperm", "Conifer"),
    plot_group = factor(plot_group, levels = c("Conifer", "Salvia", "Other Angiosperm"))
  ) %>%
  # Sorting by desc(plot_group) puts "Other Angiosperm" first in the dataframe
  arrange(desc(plot_group))
# Panel A: In-Sample Plot
plot_insample_pub <- ggplot(df_mod, aes(x = predicted_ilfm, y = ilfm)) +
  # Map the fill to our new plot_group
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_insample, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (In-Sample)",
    y = "Observed iLFM",
    fill = "Plant Group" # Clean legend title
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = "none" # Hide legend on the first plot so it doesn't duplicate
  )

# Panel B: Cross-Validation Plot
plot_cv_pub <- ggplot(df_mod, aes(x = cv_predicted_ilfm, y = ilfm)) +
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_cv, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (Unseen Species)",
    y = "Observed iLFM",
    fill = "Plant Group"
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) 
  )

# ---------------------------------------------------------
# 5. Combine and Save
# ---------------------------------------------------------
# Stitch together with patchwork and add A/B tags
# collect_guides() ensures we only have one clean legend for the whole figure
combined_plot <- plot_insample_pub + plot_cv_pub + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = "collect") 

combined_plot

ggsave(
  filename = here("results_final", "SI", "Fig_10_angiosperms_only.png"), 
  plot     = combined_plot,
  width    = 13,   # Added an extra inch of width to make room for the legend
  height   = 6,   
  units    = "in",
  dpi      = 300
)
```



