---
title: "bayesian analyses"
author: "Indra Boving"
date: "2026-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in data: 

```{r data_setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(here)
library(janitor)
source(here("scripts", "figure_info.R"))
library(ggpmisc)
```

```{r load_data}
df1 <- read_csv(here("data", "df1.csv"))
```

###### Bayesian forest plot: 
```{r model_run}
library(brms)
library(tidybayes)
library(dplyr)
library(ggplot2)
library(readr)
library(forcats)
library(posterior)

df1 <- df1 %>%
  mutate(
    species = as.factor(species),
    study   = as.factor(study),
    mwp_z   = as.numeric(scale(mwp))
  )

priors <- c(
  prior(normal(0, 1), class = "b"),          
  prior(student_t(3, 0, 1), class = "sigma") 
)

fit_spp_slopes <- brm(
  formula = ilfm ~ 0 + species + 0 + mwp_z:species + (1 | study),
  data    = df1,
  family  = gaussian(),
  prior   = priors,
  chains  = 4, cores = 4, iter = 4000, warmup = 1000,
  seed    = 123,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

pp_check(fit_spp_slopes)
summary(fit_spp_slopes)
```

# FIGURE 4. 

Figure 4. Bayesian ridgeline plot of species‐specific relationships between LFM and water potential. Posterior distributions of iLFM and z-scored midday water potential. Posterior distributions are colored based on whether they explicitly overlap with the angiosperm (non-conifer) group slope, the gymnosperm (conifer) group slope, or both. Species codes are colored by genus on the y-axis. Slope = 0 shown with a dashed vertical line. 



```{r figure_4_generation}
library(tidybayes)
library(dplyr)
library(ggplot2)
library(ggridges)
library(forcats)
library(stringr)
library(here)

SpCodes <- read_csv(here("data", "SpCodes.csv")) %>%
  mutate(species = SpCode6) %>% 
  mutate(family = familyMatch(splist = c(Genus), taxon = "plant")$family)

# ----------------------------
# 1) Extract posterior draws for species-specific slopes
# ----------------------------
slope_draws <- fit_spp_slopes %>%
  gather_draws(`^b_species.*:mwp_z$`, regex = TRUE) %>%
  mutate(
    species = gsub("^b_species", "", .variable),
    species = gsub(":mwp_z$", "", species)
  ) %>%
  transmute(.draw, species, slope = .value)

# ----------------------------
# 2) Calculate Functional Group Reference Slopes
# ----------------------------
exclude_spp <- c("ABICON", "PINJEF", "CALDEC", "PSEMEN")

# Angiosperm (Non-conifer) group slope
grand_angio <- slope_draws %>%
  filter(!species %in% exclude_spp) %>%
  group_by(.draw) %>%
  summarise(grand_slope_angio = mean(slope, na.rm = TRUE), .groups = "drop")

angio_med <- median(grand_angio$grand_slope_angio)
angio_ci  <- quantile(grand_angio$grand_slope_angio, probs = c(0.025, 0.975), na.rm = TRUE)

# Gymnosperm (Conifer-only) group slope
grand_gymno <- slope_draws %>%
  filter(species %in% exclude_spp) %>%
  group_by(.draw) %>%
  summarise(grand_slope_gymno = mean(slope, na.rm = TRUE), .groups = "drop")

gymno_med <- median(grand_gymno$grand_slope_gymno)
gymno_ci  <- quantile(grand_gymno$grand_slope_gymno, probs = c(0.025, 0.975), na.rm = TRUE)

# ----------------------------
# 3) Determine Species Overlaps & Calculate Means for Ordering
# ----------------------------
species_summary <- slope_draws %>%
  group_by(species) %>%
  summarise(
    mean = mean(slope),
    lo = quantile(slope, 0.025, na.rm = TRUE),
    hi = quantile(slope, 0.975, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    overlaps_angio = !(hi < angio_ci[[1]] | lo > angio_ci[[2]]),
    overlaps_gymno = !(hi < gymno_ci[[1]] | lo > gymno_ci[[2]]),
    overlap_cat = case_when(
      overlaps_angio & overlaps_gymno ~ "Overlaps Both",
      overlaps_angio                  ~ "Overlaps Angio.",
      overlaps_gymno                  ~ "Overlaps Gymno.",
      TRUE                            ~ "Differs (Both)"
    )
  )

# ----------------------------
# 4) Species ordering: by posterior MEAN
# ----------------------------
species_order <- species_summary %>%
  arrange(desc(mean)) %>%
  pull(species)

species_levels_for_plot <- rev(species_order)

# ----------------------------
# 5) Build plotting data frame
# ----------------------------
plot_df <- slope_draws %>%
  left_join(species_summary %>% select(species, overlap_cat), by = "species") %>%
  mutate(
    species = factor(species, levels = species_levels_for_plot),
    overlap_cat = factor(overlap_cat, levels = c(
      "Overlaps Angio.", 
      "Overlaps Gymno.", 
      "Overlaps Both", 
      "Differs (Both)"
    ))
  )

# ----------------------------
# 6) Axis text colors by Functional Group (Angio vs Gymno)
# ----------------------------
# Colors chosen to perfectly match your reference vlines
functional_colors <- c(
  "Angiosperm" = "goldenrod",
  "Gymnosperm" = "#17154f"
)

species_levels <- levels(plot_df$species)

axis_y_cols <- tibble(species = species_levels) %>%
  mutate(
    functional_group = if_else(species %in% exclude_spp, "Gymnosperm", "Angiosperm"),
    col = functional_colors[functional_group]
  ) %>%
  pull(col)

# ----------------------------
# 7) Single Panel Ridgeline Plot
# ----------------------------
xlims <- c(-0.55, 0.3)

p_slopes_final <- ggplot(plot_df, aes(x = slope, y = species, fill = overlap_cat)) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.4) +
  # Angiosperm reference line
  geom_vline(xintercept = angio_med, linewidth = 0.8, color = "goldenrod") +
  # Gymnosperm reference line
  geom_vline(xintercept = gymno_med, linewidth = 0.8, color = "#17154f") +
  geom_density_ridges(scale = 2.2, rel_min_height = 0.01, alpha = 0.85, color = NA) +
  scale_fill_manual(values = c(
    "Overlaps Angio." = "#f5bb50", 
    "Overlaps Gymno." = "#17154f", 
    "Overlaps Both"   = "grey60", 
    "Differs (Both)"  = "grey30"
  )) +
  coord_cartesian(xlim = xlims) +
  labs(
    x = expression("Posterior of species slope (iLFM ~ " * italic(Psi)[MD] * ")"),
    y = NULL,
    fill = "Group Overlap"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = c(0.86, 0.20),
    legend.direction = "vertical",
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text.y = element_text(color = axis_y_cols, face = "bold", size = 11),
    plot.margin = margin(5, 5, 5, 5)
  )

p_slopes_final

ggsave(
  filename = here("figures", "species_slope_posteriors_functional_groups.png"),
  plot     = p_slopes_final,
  width    = 7,
  height   = 6,
  units    = "in",
  dpi      = 300
)
```