---
title: "Q2"
author: "Indra Boving"
date: "2026-01-23"
output: html_document
---
```{r}
library(janitor)
library(tidyverse)
library(here)
library(U.Taxonstand) 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here('scripts', 'lfm_functions.R'))
```

Question: Across species, what are the contributions of initial LFM and minimum WP in explaining variation in minimum LFM observed at the end of the dry season (following the conceptual model in Figure 2)? 


```{r, minwpvsmaxlfm}
concept_params_df <- read_csv(here::here("data", "concept_df.csv")) %>% 
  clean_names() %>% 
  group_by(species) %>% 
  mutate(minlfm = min(lfm, na.rm = T),
         maxlfm = max(lfm, na.rm = T),
         minwp = min(mwp, na.rm = T), 
         maxwp = max(mwp, na.rm = T),
         minilfm = 1/minlfm, 
         intercept_lfm = 1/intercept,
         maxlfm_minwp = -1*minwp + maxlfm
         ) %>% 
  merge(read_csv(here("data", "sp_info.csv"))) %>% 
  clean_names()
```

#Q2: 

```{r}
color_species <- scale_color_manual(
  values = c(
    ABICON = "#17154f",
    ADEFAS = "#2f357c",
    ARBMEN = "#6c5d9e",
    
    ARCGLA = "#9d9cd5",
    ARCPAT = "#b0799a",
    CALDEC = "#59385c",
    CEACUN = "#f6b3b0",
    
    CEAPAR = "#b38711",
    CEASPI = "#e48171",
    CERBET = "#d8b847",
    
    CEACOR = "#f6b3b0",
    
    ERIFAS = "#bf3729",
    HETARB = "#e69b00",
    
    MALLAU = "#f5bb50",
    PINJEF = "#ada43b",
    PSEMEN = "#355828",
    
    QUEAGR = "#5b859e",
    QUEBER = "#1e395f",
    QUEDOU = "#75884b",
    
    QUEDUR = "#1e5a46",
    QUEGAR = "#df8d71",
    QUEKEL = "#af4f2f",
    
    SALLEU = "#d48f90",
    SALMEL = "#732f30",
    UMBCAL = "#d8b847"
  ),
  drop = FALSE
)
```


```{r, minwpvsmaxlfm}
concept_params_df <- read_csv(here::here("data", "concept_df.csv")) %>% 
  clean_names() %>% 
  group_by(species) %>% 
  mutate(minlfm = min(lfm, na.rm = T),
         maxlfm = max(lfm, na.rm = T),
         minwp = min(mwp, na.rm = T), 
         maxwp = max(mwp, na.rm = T),
         minilfm = 1/minlfm, 
         intercept_lfm = 1/intercept,
         maxlfm_minwp = -1*minwp + maxlfm
         ) %>% 
  merge(read_csv(here("data", "sp_info.csv"))) %>% 
  clean_names()
```

```{r}
SpCodes <- read_csv(here("data", "SpCodes.csv")) %>%
  mutate(species = SpCode6) %>% 
  mutate(family = familyMatch(splist = c(Genus), taxon = "plant")$family)
```

#Figure 7*:

```{r}
make_lm_plot <- function(df, xvar, yvar, xlab, ylab,
                         eq_x = NULL, eq_y = NULL, eq_col = "black") {

  model <- lm(reformulate(xvar, yvar), data = df)
  eqn <- paste0(
    "y = ", round(coef(model)[2], 3), "x + ", round(coef(model)[1], 3),
    "\nR² = ", round(summary(model)$r.squared, 3),
    "\np = ", signif(summary(model)$coefficients[2, 4], 3)
  )

  # defaults: upper-left-ish in data space
  if (is.null(eq_x)) eq_x <- min(df[[xvar]], na.rm = TRUE)
  if (is.null(eq_y)) eq_y <- max(df[[yvar]], na.rm = TRUE)

 ggplot(df, aes(
  x = .data[[xvar]],
  y = .data[[yvar]]*100,
  color = species
)) +

    geom_point(size = 2, aes(shape = functional_type)) +
    geom_smooth(method = "lm", se = FALSE, color = eq_col) +
    annotate(
      "text", x = eq_x, y = eq_y, label = eqn,
      hjust = 0, vjust = 1, size = 3.5,
      color = eq_col, fontface = "italic"
    ) +
    color_species +
   scale_shape_manual(
  values = c(
    "Angiosperm" = 16,  # filled circle
    "Conifer" = 17   # filled triangle
  )
) +
    labs(x = xlab, 
         y = ylab, 
         color = "Species", 
         shape = "Plant Type") +
    theme(legend.position = "right") +
   theme(
  legend.text  = element_text(size = 8),
  legend.title = element_text(size = 9),
  legend.key.size = unit(0.5, "lines") +
    theme(
  plot.margin = margin(5, 5, 5, 5)  # top, right, bottom, left (pt)
)
)
}

library(dplyr)
library(cowplot)

df_plot <- concept_params_df %>% 
  filter(group == "species") %>%
  mutate(
    functional_type = case_when(
      functional_type %in% c("gymnosperm") ~ "Conifer", 
      TRUE ~ as.character("Angiosperm")))
    

# Shared y-limits across both panels (Minimum LFM)
ylims_minlfm <- range(df_plot$minlfm*100, na.rm = TRUE)

# --- Panel A: Minimum LFM vs Minimum WP ---
fig4f_yminlfm <- make_lm_plot(
  df_plot,
  xvar = "minwp",
  yvar = "minlfm",
  xlab = expression(Psi[MDmin]~"(MPa)"),
  ylab = "LFMmin (%)"
) +
  coord_cartesian(ylim = ylims_minlfm)

# --- Panel B: Minimum LFM vs Model Slope ---
fig4g_yminlfm <- make_lm_plot(
  df_plot,
  xvar = "slope",
  yvar = "minlfm",
  xlab = "Slope",
  ylab = "LFMmin (%)"
) +
  coord_cartesian(ylim = ylims_minlfm)

# Legend (same as before)
legend <- get_legend(
  fig4f_yminlfm + theme(legend.box.margin = margin(0, 0, 5, 0)) +
    theme(
      legend.position = "right",
      legend.key = element_blank(),
      legend.background = element_blank(),
      legend.box.background = element_blank()
    )
)

# Combine with shared y scale + single y-axis labeling
plots_fig4a <- plot_grid(
  fig4f_yminlfm +
    theme(
      legend.position = "none",
      plot.margin = margin(5, 2, 5, 5)
    ),
  fig4g_yminlfm +
    theme(
      legend.position = "none",
      axis.title.y = element_blank(),
      axis.text.y  = element_blank(),
      axis.ticks.y = element_blank(),
      plot.margin  = margin(5, 5, 5, 2)
    ),
  nrow = 1,
  labels = LETTERS[1:2],
  label_size = 14,
  label_x = 0.02,   # move A/B inside panel
  label_y = 0.96,
  hjust = 0,
  vjust = 1,
  rel_widths = c(1, .85)
)

plots_fig4a_final <- plot_grid(plots_fig4a, legend, rel_widths = c(1, .25), nrow = 1)
plots_fig4a_final

ggsave(
  plot = plots_fig4a_final,
  filename = here::here("figures", "fig7.jpg"),
  dpi = 600, 
  width = 7, 
  height = 3
)

```
Stats for the above: 
```{r}
#Slope and minLFM
spp_df <- df_plot

# ---- 1) Clean + create percent minLFM (optional) ----
dat <- spp_df %>%
  filter(is.finite(slope), is.finite(minlfm)) %>%
  mutate(
    minlfm_pct = minlfm * 100
  )

# ---- 2) Pearson + Spearman correlations ----
pear <- cor.test(dat$slope, dat$minlfm, method = "pearson")
spear <- cor.test(dat$slope, dat$minlfm, method = "spearman", exact = FALSE)

pear
spear

# Helper to print a clean one-liner you can paste into text
cat(
  sprintf(
    "\nPearson r = %.3f (95%% CI %.3f to %.3f), p = %.3g, n = %d\n",
    unname(pear$estimate), pear$conf.int[1], pear$conf.int[2], pear$p.value, nrow(dat)
  )
)

cat(
  sprintf(
    "Spearman rho = %.3f, p = %.3g, n = %d\n\n",
    unname(spear$estimate), spear$p.value, nrow(dat)
  )
)
# ---- 3) Simple linear regression + R2 ----
m_lm <- lm(minlfm ~ slope, data = dat)
summary(m_lm)

r2 <- summary(m_lm)$r.squared
cat(sprintf("Linear model: minlfm ~ slope, R^2 = %.3f\n\n", r2))

# Also tidy output (nice for tables)
lm_tidy <- broom::tidy(m_lm, conf.int = TRUE)
lm_glance <- broom::glance(m_lm)
lm_tidy
lm_glance

# Optional: regression using percent on y-axis (same p-values, different scale)
m_lm_pct <- lm(minlfm_pct ~ slope, data = dat)
summary(m_lm_pct)
```

```{r}
#Minmpa and minLFM
spp_df <- df_plot

# ---- 1) Clean + create percent minLFM (optional) ----
dat <- spp_df %>%
  filter(is.finite(minwp), is.finite(minlfm)) %>%
  mutate(
    minlfm_pct = minlfm * 100
  )

# ---- 2) Pearson + Spearman correlations ----
pear <- cor.test(dat$minwp, dat$minlfm, method = "pearson")
spear <- cor.test(dat$minwp, dat$minlfm, method = "spearman", exact = FALSE)

pear
spear

# Helper to print a clean one-liner you can paste into text
cat(
  sprintf(
    "\nPearson r = %.3f (95%% CI %.3f to %.3f), p = %.3g, n = %d\n",
    unname(pear$estimate), pear$conf.int[1], pear$conf.int[2], pear$p.value, nrow(dat)
  )
)

cat(
  sprintf(
    "Spearman rho = %.3f, p = %.3g, n = %d\n\n",
    unname(spear$estimate), spear$p.value, nrow(dat)
  )
)
# ---- 3) Simple linear regression + R2 ----
m_lm <- lm(minlfm ~ minwp, data = dat)
summary(m_lm)

r2 <- summary(m_lm)$r.squared
cat(sprintf("Linear model: minlfm ~ minwp, R^2 = %.3f\n\n", r2))

# Also tidy output (nice for tables)
lm_tidy <- broom::tidy(m_lm, conf.int = TRUE)
lm_glance <- broom::glance(m_lm)
lm_tidy
lm_glance

# Optional: regression using percent on y-axis (same p-values, different scale)
m_lm_pct <- lm(minlfm_pct ~ minwp, data = dat)
summary(m_lm_pct)
```


#Fugure S9*: model parameter correlations: 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(here)

# ============================
# Compare model parameters only
#   - makes a params × params correlation heatmap
#   - runs once per dataset (with/without salvias)
# ============================

# ----------------------------
# Pick vars (model params only)
# ----------------------------
model_params <- c("minlfm", "maxlfm", "intercept_lfm", "slope", "minwp")

# ----------------------------
# Nice labels (edit however you want)
# ----------------------------
nice_names <- c(
  minlfm       = "LFMmin",
  maxlfm       = "LFMmax",
  intercept_lfm = "Intercept (LFM)",
  slope         = "Slope",
  minwp        = "Min Ψmd"
)

# ----------------------------
# Helper: pairwise cor.test p-values (handles NAs per pair)
# ----------------------------
pairwise_cor_p <- function(df, vars, method = "pearson") {
  pairs <- t(combn(vars, 2))
  out <- tibble(
    x = pairs[, 1],
    y = pairs[, 2]
  ) %>%
    rowwise() %>%
    mutate(
      n = sum(is.finite(df[[x]]) & is.finite(df[[y]])),
      test = list({
        xx <- df[[x]]
        yy <- df[[y]]
        ok <- is.finite(xx) & is.finite(yy)
        if (sum(ok) < 3) return(NULL)
        suppressWarnings(cor.test(xx[ok], yy[ok], method = method))
      }),
      r = if (is.null(test)) NA_real_ else unname(test$estimate),
      p = if (is.null(test)) NA_real_ else test$p.value
    ) %>%
    ungroup()

  # mirror to full matrix (x,y) and (y,x), plus diagonal
  diag_df <- tibble(x = vars, y = vars, n = NA_integer_, r = 1, p = NA_real_)
  bind_rows(
    out,
    out %>% transmute(x = y, y = x, n = n, r = r, p = p),
    diag_df
  )
}

# ----------------------------
# Make one plot function so you don't repeat code
# ----------------------------
plot_param_cor <- function(dat, outfile, method = "pearson") {

  plot_df <- dat %>%
    select(all_of(model_params)) %>%
    mutate(across(everything(), as.numeric)) %>%
    { pairwise_cor_p(., vars = model_params, method = method) } %>%
    mutate(
      x_lab = nice_names[x],
      y_lab = nice_names[y],
      stars = case_when(
        is.na(p) ~ "",
        p < 0.001 ~ "***",
        p < 0.01  ~ "**",
        p < 0.05  ~ "*",
        TRUE ~ ""
      ),
      label = case_when(
        x == y ~ "r=1",
        is.na(p) ~ "NA",
        TRUE ~ paste0("r=", round(r, 2), "\n",
                      "p=", format.pval(p, digits = 2, eps = 1e-3), "\n", stars)
      )
    ) %>%
    mutate(
      x_lab = factor(x_lab, levels = nice_names[model_params]),
      # keep same ordering on y, but reverse so first param is top row
      y_lab = factor(y_lab, levels = rev(nice_names[model_params]))
    )

  p <- ggplot(plot_df, aes(x = x_lab, y = y_lab, fill = r)) +
    geom_tile(color = "white", linewidth = 0.4) +
    geom_text(aes(label = label), size = 3, lineheight = 0.9) +
    scale_fill_gradient2(
      low = "#2c2c2c", mid = "white", high = "#b22222",
      limits = c(-1, 1), oob = squish, name = "Pearson r"
    ) +
    labs(x = NULL, y = NULL) +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
      legend.key.height = unit(1.2, "lines")
    )

  ggsave(
    filename = outfile,
    plot = p,
    width = 5.5,
    height = 5,
    units = "in"
  )

  p
}

# ----------------------------
# Run it: with salvias
# ----------------------------
p_params_with <- plot_param_cor(
  dat = df_plot,
  outfile = here("figures", "modelparams_correlation_with_salvias.jpeg")
)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(here)
library(GGally)

# ----------------------------
# Pick vars (model params only)
# ----------------------------
model_params <- c("minlfm", "maxlfm", "intercept_lfm", "slope", "minwp")

# ----------------------------
# Nice labels
# ----------------------------
nice_names <- c(
  minlfm        = "LFMmin",
  maxlfm        = "LFMmax",
  intercept_lfm = "Intercept (LFM)",
  slope         = "Slope",
  minwp         = "Min Ψmd"
)

# ----------------------------
# One plot: ggpairs
#   - lower: correlation tiles (with r and p)
#   - upper: scatterplots
# ----------------------------
plot_param_pairs <- function(dat, outfile, method = "pearson") {

  method_use <- method  # capture once to avoid recursive default arg issues

  dat_use <- dat %>%
    select(all_of(model_params)) %>%
    mutate(across(everything(), as.numeric)) %>%
    rename_with(~ nice_names[.x], all_of(model_params))  # prettier labels

  # ---- lower triangle: correlation tile ----
  cor_tile <- function(data, mapping, ...) {
    x <- GGally::eval_data_col(data, mapping$x)
    y <- GGally::eval_data_col(data, mapping$y)

    ok <- is.finite(x) & is.finite(y)
    if (sum(ok) < 3) {
      return(
        ggplot() +
          theme_void() +
          annotate("text", x = 0.5, y = 0.5, label = "NA")
      )
    }

    ct <- suppressWarnings(cor.test(x[ok], y[ok], method = method_use))
    r  <- unname(ct$estimate)
    p  <- ct$p.value

    stars <- dplyr::case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE ~ ""
    )

    label <- paste0(
      "r=", round(r, 2), "\n",
      "p=", format.pval(p, digits = 2, eps = 1e-3), "\n", stars
    )

    ggplot() +
      geom_tile(aes(x = 1, y = 1, fill = r), color = "white", linewidth = 0.4) +
      geom_text(aes(x = 1, y = 1, label = label), size = 3, lineheight = 0.9) +
      scale_fill_gradient2(
        low = "#2c2c2c", mid = "white", high = "#b22222",
        limits = c(-1, 1), oob = squish
      ) +
      theme_void() +
      theme(legend.position = "none")
  }

  # ---- upper triangle: scatter ----
  scat_panel <- function(data, mapping, ...) {
    ggplot(data = data, mapping = mapping) +
      geom_point(alpha = 0.7, size = 1) +
      theme(
        panel.grid = element_blank(),
        axis.title = element_blank()
      )
  }

  # ---- diagonal: density (or swap to points if you prefer) ----
  diag_density <- function(data, mapping, ...) {
    ggplot(data = data, mapping = mapping) +
      geom_density(linewidth = 0.6, adjust = 1.2) +
      theme(
        panel.grid = element_blank(),
        axis.title = element_blank()
      )
  }

  g <- GGally::ggpairs(
    dat_use,
    upper = list(continuous = scat_panel),
    lower = list(continuous = cor_tile),
    diag  = list(continuous = diag_density)
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(size = 10),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "right"
    )

  ggsave(outfile, g, width = 7.5, height = 7.5, units = "in")
  g
}

# ----------------------------
# Run once
# ----------------------------
p_params <- plot_param_pairs(
  dat = df_plot,
  outfile = here("figures", "modelparams_pairs_scatter_corr.jpeg")
)

```


####Figure 5. (isocline)

Using common slope for all species: 

```{r, fig.height=4, fig.width=6}
# Read species results
# spres_old <- read.csv(here('results', 'species-results.csv'))

# Ensure 'spres' has all the necessary columns from the previous steps
spres <- concept_params_df %>% 
  filter(group == "grand") %>% 
  mutate(model1.int = intercept,
         model1.slp = slope,  # Ensure slope is named 'model1.slp'
         SpCode6 = species, 
         minWP = minwp)

# Inspect the modified dataframe
head(spres)

# Generate a sequence of inverse LFM saturation values
ilfm.sat.vec <- seq(0.25, 1.25, length.out = 31)

# Initialize an empty list to store critical MWP values for each species
MWPcrit_list <- list()

# Loop through each unique species and calculate the critical MWP using species-specific slopes
for (sp in unique(spres$species)) {
  
  # Get the species-specific slope from 'spres'
  species_slope <- spres %>% 
    filter(species == sp) %>% 
    pull(model1.slp)
  
  # Initialize a vector to hold critical MWP values for the current species
  MWPcrit <- c()
  
  # Calculate critical MWP for the current species
  for (i in 1:length(ilfm.sat.vec)) {
    MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.7)  # Use species-specific slope
  }
  
  # Store the results for each species in the list
  MWPcrit_list[[sp]] <- MWPcrit
}

# Convert inverse LFM saturation values back to LFM
lfm.sat.vec <- 1 / ilfm.sat.vec

# Step 7: Plot the conceptual model curve
plot(MWPcrit, lfm.sat.vec, 
     xlim = c(-8, 0), ylim = c(0, 4),
     type = 'l', lwd = 0, col = "blue",
     xlab = 'Minimum Midday Water Potential (MPa)', 
     ylab = 'LFM @ Midday Water Potential = 0')

# Step 8: Repeat the calculations for different LFM thresholds

# 60% LFM
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.62)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "maroon")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "60% LFM", pos = 4, col = "maroon", cex = 0.8)

# 77% LFM
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.77)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "goldenrod")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.8)

# 88% LFM
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.88)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "forestgreen")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "88% LFM", pos = 2, col = "forestgreen", cex = 0.8)

# Step 9: Add species data points to the conceptual model plot
# text(I(1 / model1.int) ~ minWP,
#      data = spres,
#      labels = spres$SpCode6,
#      pos = 4,
#      cex = 0.6) # Reduced text size to prevent overlap

# Assign unique letters to unique species
unique_species <- unique(spres$SpCode6)
species_letters <- setNames(letters[1:length(unique_species)], unique_species)

# Add a new column in 'spres' mapping species to their assigned letters
spres$Letter <- species_letters[spres$SpCode6]

# Create a letter-to-species lookup table
letter_mapping <- data.frame(
  Letter = spres$Letter,
  Species = spres$SpCode6
) %>% distinct()  # Ensure each species appears only once

# Define color palette for species
color_species = c(
    ABICON = "#17154f",
    ADEFAS = "#2f357c",
    ARBMEN = "#6c5d9e",
    
    ARCGLA = "#9d9cd5",
    ARCPAT = "#b0799a",
    CALDEC = "#59385c",
    CEACUN = "#f6b3b0",
    
    CEAPAR = "#b38711",
    CEASPI = "#e48171",
    CERBET = "#d8b847",
    
    ERIFAS = "#bf3729",
    HETARB = "#e69b00",
    
    MALLAU = "#f5bb50",
    PINJEF = "#ada43b",
    PSEMEN = "#355828",
    
    QUEAGR = "#5b859e",
    QUEBER = "#1e395f",
    QUEDOU = "#75884b",
    
    QUEDUR = "#1e5a46",
    QUEGAR = "#df8d71",
    QUEKEL = "#af4f2f",
    
    SALLEU = "#d48f90",
    SALMEL = "#732f30",
    UMBCAL = "#d8b847"
  )

species_cols <- color_species #so it works with isocline 
# Ensure we have enough colors for all unique species
color_species <- setNames(color_species[1:length(unique_species)], unique_species)

# Add species-specific labels at the correct locations
text(spres$minWP, 1 / spres$model1.int,  # (x, y) coordinates
     labels = spres$Letter,                # Assign unique letters instead of species codes
     col = color_species[spres$SpCode6],  # Assign colors by species
     #pos = 5,                               # Adjust text positioning
     cex = 1)                             # Adjust text size to avoid overlap

```
To save the isocline: 
```{r}
# Save base R plot as high-resolution JPEG
jpeg(filename = here::here("figures", "fig4.jpg"), 
     width = 5, height = 4, units = "in", res = 600)

# Reduce whitespace: bottom, left, top, right margins
par(mar = c(3.5, 3.5, 1, 1), mgp = c(2, 0.6, 0))  # mgp controls label spacing
# Re-run the plotting code below to draw inside the jpeg device

# Step 7: Plot the conceptual model curve
plot(MWPcrit, lfm.sat.vec, 
     xlim = c(-8, 0), ylim = c(0, 4),
     type = 'l', lwd = 0, col = "blue",
     xlab = 'Minimum Midday Water Potential (MPa)', 
     ylab = 'LFM @ Midday Water Potential = 0')

# 60% LFM line
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.62)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "maroon")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "60% LFM", pos = 4, col = "maroon", cex = 0.8)

# 77% LFM line
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.77)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "goldenrod")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.8)

# 88% LFM line
for (i in 1:length(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.88)
}
lfm.sat.vec <- 1 / ilfm.sat.vec
lines(MWPcrit, lfm.sat.vec, lwd = 1, col = "forestgreen")
text(MWPcrit[length(MWPcrit)], lfm.sat.vec[length(lfm.sat.vec)], labels = "88% LFM", pos = 2, col = "forestgreen", cex = 0.8)

# Add species labels (letters)
text(spres$minWP, 1 / spres$model1.int,
     labels = spres$Letter,
     col = color_species[spres$SpCode6],
     cex = 1)

# Close the jpeg device
dev.off()

# Print letter-to-species mapping table
print(letter_mapping)

# Load necessary package
library(knitr)

# Print the letter-to-species mapping as a formatted table
letter_mapping %>% 
  kable(format = "pipe", col.names = c("Letter", "Species"))

write.csv(letter_mapping, file = here::here("figures", "letter_mapping.csv"), row.names = FALSE)
```
To save isocline labels: 

```{r}
library(dplyr)
library(knitr)
library(kableExtra)
library(here)

# Create the colored table
letter_mapping_colored <- letter_mapping %>%
  mutate(Letter = paste0("<span style='color:", color_species[Species], "'>", Letter, "</span>"))

# Generate the HTML table
table_html <- letter_mapping_colored %>%
  kable("html", escape = FALSE, col.names = c("Letter", "Species")) %>%
  kable_styling(full_width = FALSE, position = "left")

# Build full HTML page with styling for 2-column layout and smaller text
html_page <- paste0(
  "<!DOCTYPE html>
  <html>
  <head>
    <meta charset='UTF-8'>
    <title>Letter Mapping</title>
    <style>
      body {
        font-size: 0.85em;
        column-count: 2;
        column-gap: 2em;
        font-family: sans-serif;
      }
      table {
        margin-bottom: 1em;
        break-inside: avoid;
      }
    </style>
  </head>
  <body>",
  table_html,
  "</body>
  </html>"
)

# Save to file
writeLines(html_page, here("figures", "letter_mapping.html"))
```

####Figure 5, (genus)
```{r}
library(dplyr)

# ----------------------------
# Build spres from parameters
# ----------------------------
spres <- concept_params_df %>% 
  filter(group == "grand") %>% 
  mutate(
    model1.int = intercept,
    model1.slp = slope,      # slope column for convenience
    SpCode6    = species, 
    minWP      = minwp
  )

# ----------------------------
# Join Genus info (REQUIRED)
#   - Replace `species_key_df` with your actual table name
#   - Must contain columns: SpCode6, Genus
# ----------------------------
spres <- spres %>%
  left_join(SpCodes %>% select(SpCode6, Genus), by = "SpCode6")

# Optional check for any species missing a Genus
if (any(is.na(spres$Genus))) {
  warning("Some SpCode6 have no Genus in species_key_df.")
}

# ----------------------------
# Generate conceptual curves
# ----------------------------
ilfm.sat.vec <- seq(0.25, 1.25, length.out = 31)

# If you truly want species-specific MWPcrit per species, your list builder is fine.
# Keeping for completeness (not used directly in the plot below):

MWPcrit_list <- list()
for (sp in unique(spres$species)) {
  species_slope <- spres %>% filter(species == sp) %>% pull(model1.slp)
  MWPcrit <- numeric(length(ilfm.sat.vec))
  for (i in seq_along(ilfm.sat.vec)) {
    MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1 / 0.7)
  }
  MWPcrit_list[[sp]] <- MWPcrit
}

# Use a "typical" slope for drawing the conceptual lines (so code below is robust).
# If you prefer a different summary (mean, a specific species, etc.), change here:
species_slope <- median(spres$model1.slp, na.rm = TRUE)

# ----------------------------
# Plot open: 
# ----------------------------

# Open JPEG device
jpeg(
  filename = here::here("figures", "Fig2_genus.jpg"),
  width    = 7,   
  height   = 4.5,     
  units    = "in",
  res      = 300    # high resolution
)


## 1) Make room on the right (set before plotting)

op <- par(no.readonly = TRUE)                  # save current par
# Tight margins + space for legend in right margin
par(mar = c(4, 4, 0.5, 6), mgp = c(2.2, 0.6, 0), xpd = TRUE) # add right margin & allow drawing in margins

# ----------------------------
# Colors by Genus (CHANGE HERE)
# ----------------------------
# Okabe–Ito palette; expand if more genera than base colors
okabe_ito <- c("#17154f", "#f6b3b0",  "#732f30", "#e69b00", "#5b859e", "#b38711", "#6c5d9e",  "#1e395f", "#e48171", "#bf3729",  "#355828", "#ab84a5","#9d9cd5", "#ada43b", "#5b859e", "#75884b", "#df8d71", "#af4f2f", "#d48f90", "#59385c", "#d8b847")

genus_levels <- sort(unique(spres$Genus))
ng <- length(genus_levels)
genus_colors <- if (ng > length(okabe_ito)) {
  grDevices::colorRampPalette(okabe_ito)(ng)
} else {
  okabe_ito[seq_len(ng)]
}
genus_cols <- setNames(genus_colors, genus_levels)
# ----------------------------
# Base curve at 70% (reference), multiplied by 100
# ----------------------------
# --- Build the base (70%) reference once, then scale y to percent for plotting
MWPcrit <- numeric(length(ilfm.sat.vec))
for (i in seq_along(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1/0.7)
}
lfm_pct <- (1 / ilfm.sat.vec) * 100  # scale to percent for plotting

plot(
  MWPcrit, lfm_pct,
  xlim = c(-8, 0.1), ylim = c(0, 400),
  type = "l", lwd = 0, col = "blue",
  xlab = "Minimum Midday Water Potential (MPa)",
  ylab = "LFM (%) @ Midday Water Potential = 0",
  yaxt = "n"
)
axis(2, at = seq(0, 400, 100), labels = seq(0, 400, 100), las = 1)

# --- 60% LFM curve (x depends on threshold; y is the same lfm_pct scale)
for (i in seq_along(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1/0.62)
}
lines(MWPcrit, lfm_pct, lwd = 1.5, col = "maroon")
text(tail(MWPcrit, 1), tail(lfm_pct, 1), labels = "60% LFM", pos = 1, col = "maroon", cex = 0.9)

# --- 77% LFM curve
for (i in seq_along(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1/0.77)
}
lines(MWPcrit, lfm_pct, lwd = 1.5, col = "goldenrod")
text(tail(MWPcrit, 1), tail(lfm_pct, 1), labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.9)

# --- 88% LFM curve
for (i in seq_along(ilfm.sat.vec)) {
  MWPcrit[i] <- solveForX(c(ilfm.sat.vec[i], species_slope), 1/0.88)
}
lines(MWPcrit, lfm_pct, lwd = 1.5, col = "forestgreen")
text(tail(MWPcrit, 1), tail(lfm_pct, 1), labels = "88% LFM", pos = 2, col = "forestgreen", cex = 0.9)

# --- Species letters (also scaled to percent on y)
text(
  x = spres$minWP,
  y = (1 / spres$model1.int) * 100,
  labels = spres$Letter,
  col = genus_cols[spres$Genus],
  cex = 1
)

# ----------------------------
# Assign unique letters per species (unchanged)
# ----------------------------
unique_species <- unique(spres$SpCode6)
species_letters <- setNames(letters[seq_along(unique_species)], unique_species)
spres$Letter <- species_letters[spres$SpCode6]

# Lookup (if needed elsewhere)
letter_mapping <- data.frame(
  Letter  = spres$Letter,
  Species = spres$SpCode6
) %>% distinct()
```


###Figure 5, points & labels

#####Species

```{r}
# ============================
# Conceptual curves + species points (base R)
#   - builds spres from concept_params_df
#   - joins Genus from SpCodes
#   - draws conceptual isoclines + species points colored by species
#   - labels selected species (1 label per species)
#   - writes a JPEG to /figures
# ============================

library(dplyr)
library(here)

# ----------------------------
# Required helper (you said you already have this)
# ----------------------------
# solveForX <- function(cs, yval = 0.7) {
#   # y = m*x + b
#   return((yval - cs[1]) / cs[2])
# }

# ----------------------------
# Species color palette (named vector)
#   NOTE: make sure CEACOR is included
# ----------------------------
color_species <- c(
  ABICON = "#17154f",
  ADEFAS = "#2f357c",
  ARBMEN = "#6c5d9e",
  ARCGLA = "#9d9cd5",
  ARCPAT = "#b0799a",
  CALDEC = "#59385c",
  CEACOR = "#f6b3b0",  # <- added (was missing)
  CEACUN = "#f6b3b0",
  CEAPAR = "#b38711",
  CEASPI = "#e48171",
  CERBET = "#d8b847",
  ERIFAS = "#bf3729",
  HETARB = "#e69b00",
  MALLAU = "#f5bb50",
  PINJEF = "#ada43b",
  PSEMEN = "#355828",
  QUEAGR = "#5b859e",
  QUEBER = "#1e395f",
  QUEDOU = "#75884b",
  QUEDUR = "#1e5a46",
  QUEGAR = "#df8d71",
  QUEKEL = "#af4f2f",
  SALLEU = "#d48f90",
  SALMEL = "#732f30",
  UMBCAL = "#d8b847"
)

# ----------------------------
# Build spres from parameters
# ----------------------------
spres <- concept_params_df %>%
  filter(group == "grand") %>%
  mutate(
    model1.int = intercept,
    model1.slp = slope,
    SpCode6    = species,
    minWP      = minwp
  ) %>%
  left_join(SpCodes %>% select(SpCode6, Genus), by = "SpCode6")

if (any(is.na(spres$Genus))) {
  warning("Some SpCode6 have no Genus in SpCodes: ",
          paste(unique(spres$SpCode6[is.na(spres$Genus)]), collapse = ", "))
}

# ----------------------------
# Species colors aligned to data
# ----------------------------
species_levels <- sort(unique(spres$SpCode6))

species_cols <- color_species[species_levels]
missing_spp <- species_levels[is.na(species_cols)]
if (length(missing_spp) > 0) {
  warning("Missing species in color_species (will be grey50): ",
          paste(missing_spp, collapse = ", "))
  species_cols[is.na(species_cols)] <- "grey50"
}
names(species_cols) <- species_levels

# ----------------------------
# Generate conceptual curves
# ----------------------------
# Standard vector for Maroon and Goldenrod lines (caps at 350% LFM)
ilfm.sat.vec <- seq(1 / 3.5, 1.25, length.out = 31)

# Custom vector for the Green line: starts at 350% LFM (1/3.5) 
# and ends exactly where x = 0 (which is 1 / 0.88)
ilfm.sat.vec.green <- seq(1 / 3.5, 1 / 0.88, length.out = 31)

# "Typical" slope for conceptual lines
species_slope <- median(spres$model1.slp, na.rm = TRUE)

# Base y scales corresponding to ilfm values
lfm_pct       <- (1 / ilfm.sat.vec) * 100
lfm_pct_green <- (1 / ilfm.sat.vec.green) * 100

make_MWPcrit <- function(thresh, ilfm_vec, slope) {
  out <- numeric(length(ilfm_vec))
  for (i in seq_along(ilfm_vec)) {
    out[i] <- solveForX(c(ilfm_vec[i], slope), 1 / thresh)
  }
  out
}

MWPcrit_60 <- make_MWPcrit(0.62, ilfm.sat.vec, species_slope)
MWPcrit_77 <- make_MWPcrit(0.77, ilfm.sat.vec, species_slope)
MWPcrit_88 <- make_MWPcrit(0.88, ilfm.sat.vec.green, species_slope) # Uses green vector

# ----------------------------
# Plot open
# ----------------------------
jpeg(
  filename = here("figures", "Fig2_species.jpg"),
  width = 7,
  height = 4.5,
  units = "in",
  res = 300
)

op <- par(no.readonly = TRUE)
par(mar = c(4, 4, 0.5, 7), mgp = c(2.2, 0.6, 0), xpd = TRUE)

# ----------------------------
# Set up plot (axes only)
# ----------------------------
plot(
  x = MWPcrit_60, y = lfm_pct,
  xlim = c(-8, 0.2), ylim = c(0, 400),
  type = "n",
  xlab = "Minimum Midday Water Potential (MPa)",
  ylab = "LFM (%) @ Midday Water Potential = 0",
  yaxt = "n"
)
axis(2, at = seq(0, 400, 50), labels = seq(0, 400, 50), las = 1)

# ----------------------------
# Draw conceptual curves
# ----------------------------
lines(MWPcrit_60, lfm_pct, lwd = 1.5, col = "maroon")
text(MWPcrit_60[1], lfm_pct[1],
     labels = "62% LFM", pos = 3, col = "maroon", cex = 0.9)

lines(MWPcrit_77, lfm_pct, lwd = 1.5, col = "goldenrod")
text(MWPcrit_77[1], lfm_pct[1],
     labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.9)

lines(MWPcrit_88, lfm_pct_green, lwd = 1.5, col = "forestgreen")
text(MWPcrit_88[1], lfm_pct_green[1],
     labels = "88% LFM", pos = 4, col = "forestgreen", cex = 0.9)

# ----------------------------
# Plot points (colored by species)
# ----------------------------
points(
  x = spres$minWP,
  y = (1 / spres$model1.int) * 100,
  pch = 16,
  cex = 1.1,
  col = species_cols[spres$SpCode6]
)

# ----------------------------
# Labels for selected species (one label per species)
# ----------------------------
label_spp <- c("PSEMEN", "UMBCAL", "ABCO", "PIJE", "PINJEF", "ABICON",
                "CALDEC")

special <- c("PIJE", "PINJEF", "CALDEC", "QUEKEL", "CEACOR")

spres_lab <- spres %>%
  filter(SpCode6 %in% label_spp) %>%
  mutate(
    x = minWP,
    y = (1 / model1.int) * 100
  ) %>%
  group_by(SpCode6) %>%
  slice(1) %>%
  ungroup()

# right-side labels
spres_lab_other <- spres_lab %>% filter(!SpCode6 %in% special)
if (nrow(spres_lab_other) > 0) {
  text(
    x = spres_lab_other$x,
    y = spres_lab_other$y,
    labels = spres_lab_other$SpCode6,
    pos = 4,
    offset = 0.3,
    cex = 0.75,
    col = species_cols[spres_lab_other$SpCode6]
  )
}

# special labels above point
spres_lab_special <- spres_lab %>% filter(SpCode6 %in% special)
if (nrow(spres_lab_special) > 0) {
  text(
    x = spres_lab_special$x,
    y = spres_lab_special$y,
    labels = spres_lab_special$SpCode6,
    pos = 3,
    offset = 0.45,
    adj = c(0.5, 0),
    cex = 0.75,
    col = species_cols[spres_lab_special$SpCode6]
  )
}

# ----------------------------
# Legend (to the right)
# ----------------------------
usr <- par("usr")   # c(xmin, xmax, ymin, ymax)
pin <- par("pin")   # plot dimensions in inches
x_per_inch <- diff(usr[1:2]) / pin[1]
x_shift <- 0.9 * x_per_inch  # ~0.9 inch to the right

legend(
  x = usr[2] + x_shift,
  y = usr[4],
  xjust = 1, yjust = 1,
  legend = species_levels,
  pch = 16,
  col = species_cols[species_levels],
  bty = "n",
  cex = 0.7,
  title = "Species"
)

par(op)
dev.off()
```

#####Colored by dec/evergreen



```{r}
# ============================
# Conceptual curves + species points (base R)
#   - builds spres from concept_params_df
#   - joins Genus from SpCodes
#   - draws conceptual isoclines + species points colored by leaf habit (dec_ever)
#   - labels selected species (1 label per species)
#   - writes a JPEG to /figures
# ============================

library(dplyr)
library(here)

# ----------------------------
# Required helper (you said you already have this)
# ----------------------------
# solveForX <- function(cs, yval = 0.7) {
#   # y = m*x + b
#   return((yval - cs[1]) / cs[2])
# }

# ----------------------------
# Leaf Habit color palette (named vector)
# ----------------------------
color_habit <- c(
  "Dec"   = "#df8d71",  # Nice terracotta red for Deciduous
  "Ever"  = "#1e5a46"   # Nice deep forest green for Evergreen
)

# ----------------------------
# Build spres from parameters
# ----------------------------
spres <- concept_params_df %>%
  filter(group == "grand") %>%
  mutate(
    model1.int = intercept,
    model1.slp = slope,
    SpCode6    = species,
    minWP      = minwp,
    dec_ever   = dec_ever  # Pulling leaf habit into our plotting dataframe
  ) %>%
  left_join(SpCodes %>% select(SpCode6, Genus), by = "SpCode6")

if (any(is.na(spres$Genus))) {
  warning("Some SpCode6 have no Genus in SpCodes: ",
          paste(unique(spres$SpCode6[is.na(spres$Genus)]), collapse = ", "))
}

# ----------------------------
# Species colors aligned to data (for text labels only)
# ----------------------------
color_species <- c(
  ABICON = "#17154f", ADEFAS = "#2f357c", ARBMEN = "#6c5d9e", ARCGLA = "#9d9cd5",
  ARCPAT = "#b0799a", CALDEC = "#59385c", CEACOR = "#f6b3b0", CEACUN = "#f6b3b0",
  CEAPAR = "#b38711", CEASPI = "#e48171", CERBET = "#d8b847", ERIFAS = "#bf3729",
  HETARB = "#e69b00", MALLAU = "#f5bb50", PINJEF = "#ada43b", PSEMEN = "#355828",
  QUEAGR = "#5b859e", QUEBER = "#1e395f", QUEDOU = "#75884b", QUEDUR = "#1e5a46",
  QUEGAR = "#df8d71", QUEKEL = "#af4f2f", SALLEU = "#d48f90", SALMEL = "#732f30",
  UMBCAL = "#d8b847"
)

species_levels <- sort(unique(spres$SpCode6))
species_cols <- color_species[species_levels]
species_cols[is.na(species_cols)] <- "grey50"
names(species_cols) <- species_levels

# ----------------------------
# Generate conceptual curves
# ----------------------------
# Standard vector for Maroon and Goldenrod lines (caps at 350% LFM)
ilfm.sat.vec <- seq(1 / 3.5, 1.25, length.out = 31)

# Custom vector for the Green line: starts at 350% LFM (1/3.5) 
# and ends exactly where x = 0 (which is 1 / 0.88)
ilfm.sat.vec.green <- seq(1 / 3.5, 1 / 0.88, length.out = 31)

# "Typical" slope for conceptual lines
species_slope <- median(spres$model1.slp, na.rm = TRUE)

# Base y scales corresponding to ilfm values
lfm_pct       <- (1 / ilfm.sat.vec) * 100
lfm_pct_green <- (1 / ilfm.sat.vec.green) * 100

make_MWPcrit <- function(thresh, ilfm_vec, slope) {
  out <- numeric(length(ilfm_vec))
  for (i in seq_along(ilfm_vec)) {
    out[i] <- solveForX(c(ilfm_vec[i], slope), 1 / thresh)
  }
  out
}

MWPcrit_60 <- make_MWPcrit(0.62, ilfm.sat.vec, species_slope)
MWPcrit_77 <- make_MWPcrit(0.77, ilfm.sat.vec, species_slope)
MWPcrit_88 <- make_MWPcrit(0.88, ilfm.sat.vec.green, species_slope) # Uses green vector

# ----------------------------
# Plot open
# ----------------------------
jpeg(
  filename = here("figures", "Fig2_species.jpg"),
  width = 7,
  height = 4.5,
  units = "in",
  res = 300
)

op <- par(no.readonly = TRUE)
par(mar = c(4, 4, 0.5, 7), mgp = c(2.2, 0.6, 0), xpd = TRUE)

# ----------------------------
# Set up plot (axes only)
# ----------------------------
plot(
  x = MWPcrit_60, y = lfm_pct,
  xlim = c(-8, 0.2), ylim = c(0, 400),
  type = "n",
  xlab = "Minimum Midday Water Potential (MPa)",
  ylab = "LFM (%) @ Midday Water Potential = 0",
  yaxt = "n"
)
axis(2, at = seq(0, 400, 50), labels = seq(0, 400, 50), las = 1)

# ----------------------------
# Draw conceptual curves
# ----------------------------
lines(MWPcrit_60, lfm_pct, lwd = 1.5, col = "maroon")
text(MWPcrit_60[1], lfm_pct[1],
     labels = "62% LFM", pos = 3, col = "maroon", cex = 0.9)

lines(MWPcrit_77, lfm_pct, lwd = 1.5, col = "goldenrod")
text(MWPcrit_77[1], lfm_pct[1],
     labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.9)

lines(MWPcrit_88, lfm_pct_green, lwd = 1.5, col = "forestgreen")
text(MWPcrit_88[1], lfm_pct_green[1],
     labels = "88% LFM", pos = 4, col = "forestgreen", cex = 0.9)

# ----------------------------
# Plot points (colored by leaf habit)
# ----------------------------
points(
  x = spres$minWP,
  y = (1 / spres$model1.int) * 100,
  pch = 16,
  cex = 1.1,
  col = color_habit[spres$dec_ever] # Updated mapping to leaf habit colors
)

# ----------------------------
# Labels for selected species (one label per species)
# ----------------------------
label_spp <- c("PSEMEN", "UMBCAL", "ABCO", "PIJE", "PINJEF", "ABICON",
               "CALDEC")

special <- c("PIJE", "PINJEF", "CALDEC", "QUEKEL", "CEACOR")

spres_lab <- spres %>%
  filter(SpCode6 %in% label_spp) %>%
  mutate(
    x = minWP,
    y = (1 / model1.int) * 100
  ) %>%
  group_by(SpCode6) %>%
  slice(1) %>%
  ungroup()

# right-side labels
spres_lab_other <- spres_lab %>% filter(!SpCode6 %in% special)
if (nrow(spres_lab_other) > 0) {
  text(
    x = spres_lab_other$x,
    y = spres_lab_other$y,
    labels = spres_lab_other$SpCode6,
    pos = 4,
    offset = 0.3,
    cex = 0.75,
    col = species_cols[spres_lab_other$SpCode6]
  )
}

# special labels above point
spres_lab_special <- spres_lab %>% filter(SpCode6 %in% special)
if (nrow(spres_lab_special) > 0) {
  text(
    x = spres_lab_special$x,
    y = spres_lab_special$y,
    labels = spres_lab_special$SpCode6,
    pos = 3,
    offset = 0.45,
    adj = c(0.5, 0),
    cex = 0.75,
    col = species_cols[spres_lab_special$SpCode6]
  )
}

# ----------------------------
# Legend (to the right, updated for Leaf Habit)
# ----------------------------
usr <- par("usr")   # c(xmin, xmax, ymin, ymax)
pin <- par("pin")   # plot dimensions in inches
x_per_inch <- diff(usr[1:2]) / pin[1]
x_shift <- 0.9 * x_per_inch  # ~0.9 inch to the right

legend(
  x = usr[2] + x_shift,
  y = usr[4],
  xjust = 1, yjust = 1,
  legend = c("Deciduous", "Evergreen"), # Clear, publication-ready labels
  pch = 16,
  col = color_habit[c("Dec", "Ever")],
  bty = "n",
  cex = 0.8,
  title = "Leaf Habit"
)

par(op)
dev.off()
```

####Growth form:


```{r}
# ============================
# Conceptual curves + species points (base R)
#   - builds spres from concept_params_df
#   - joins Genus from SpCodes
#   - draws conceptual isoclines + species points colored by growth form
#   - labels selected species (1 label per species)
#   - writes a JPEG to /figures
# ============================

library(dplyr)
library(here)

# ----------------------------
# Required helper (you said you already have this)
# ----------------------------
# solveForX <- function(cs, yval = 0.7) {
#   # y = m*x + b
#   return((yval - cs[1]) / cs[2])
# }

# ----------------------------
# Growth Form color palette (named vector)
# ----------------------------
color_growth <- c(
  "tree"  = "#1e395f",  # Deep slate blue for trees
  "shrub" = "#d8b847"   # Warm golden-yellow for shrubs
)

# ----------------------------
# Build spres from parameters
# ----------------------------
spres <- concept_params_df %>%
  filter(group == "grand") %>%
  mutate(
    model1.int  = intercept,
    model1.slp  = slope,
    SpCode6     = species,
    minWP       = minwp,
    growth_form = growth_form  # Pulling growth form into our plotting dataframe
  ) %>%
  left_join(SpCodes %>% select(SpCode6, Genus), by = "SpCode6")

if (any(is.na(spres$Genus))) {
  warning("Some SpCode6 have no Genus in SpCodes: ",
          paste(unique(spres$SpCode6[is.na(spres$Genus)]), collapse = ", "))
}

# ----------------------------
# Species colors aligned to data (for text labels only)
# ----------------------------
color_species <- c(
  ABICON = "#17154f", ADEFAS = "#2f357c", ARBMEN = "#6c5d9e", ARCGLA = "#9d9cd5",
  ARCPAT = "#b0799a", CALDEC = "#59385c", CEACOR = "#f6b3b0", CEACUN = "#f6b3b0",
  CEAPAR = "#b38711", CEASPI = "#e48171", CERBET = "#d8b847", ERIFAS = "#bf3729",
  HETARB = "#e69b00", MALLAU = "#f5bb50", PINJEF = "#ada43b", PSEMEN = "#355828",
  QUEAGR = "#5b859e", QUEBER = "#1e395f", QUEDOU = "#75884b", QUEDUR = "#1e5a46",
  QUEGAR = "#df8d71", QUEKEL = "#af4f2f", SALLEU = "#d48f90", SALMEL = "#732f30",
  UMBCAL = "#d8b847"
)

species_levels <- sort(unique(spres$SpCode6))
species_cols <- color_species[species_levels]
species_cols[is.na(species_cols)] <- "grey50"
names(species_cols) <- species_levels

# ----------------------------
# Generate conceptual curves
# ----------------------------
# Standard vector for Maroon and Goldenrod lines (caps at 350% LFM)
ilfm.sat.vec <- seq(1 / 3.5, 1.25, length.out = 31)

# Custom vector for the Green line: starts at 350% LFM (1/3.5) 
# and ends exactly where x = 0 (which is 1 / 0.88)
ilfm.sat.vec.green <- seq(1 / 3.5, 1 / 0.88, length.out = 31)

# "Typical" slope for conceptual lines
species_slope <- median(spres$model1.slp, na.rm = TRUE)

# Base y scales corresponding to ilfm values
lfm_pct       <- (1 / ilfm.sat.vec) * 100
lfm_pct_green <- (1 / ilfm.sat.vec.green) * 100

make_MWPcrit <- function(thresh, ilfm_vec, slope) {
  out <- numeric(length(ilfm_vec))
  for (i in seq_along(ilfm_vec)) {
    out[i] <- solveForX(c(ilfm_vec[i], slope), 1 / thresh)
  }
  out
}

MWPcrit_60 <- make_MWPcrit(0.62, ilfm.sat.vec, species_slope)
MWPcrit_77 <- make_MWPcrit(0.77, ilfm.sat.vec, species_slope)
MWPcrit_88 <- make_MWPcrit(0.88, ilfm.sat.vec.green, species_slope) # Uses green vector

# ----------------------------
# Plot open
# ----------------------------
jpeg(
  filename = here("figures", "Fig2_species.jpg"),
  width = 7,
  height = 4.5,
  units = "in",
  res = 300
)

op <- par(no.readonly = TRUE)
par(mar = c(4, 4, 0.5, 7), mgp = c(2.2, 0.6, 0), xpd = TRUE)

# ----------------------------
# Set up plot (axes only)
# ----------------------------
plot(
  x = MWPcrit_60, y = lfm_pct,
  xlim = c(-8, 0.2), ylim = c(0, 400),
  type = "n",
  xlab = "Minimum Midday Water Potential (MPa)",
  ylab = "LFM (%) @ Midday Water Potential = 0",
  yaxt = "n"
)
axis(2, at = seq(0, 400, 50), labels = seq(0, 400, 50), las = 1)

# ----------------------------
# Draw conceptual curves
# ----------------------------
lines(MWPcrit_60, lfm_pct, lwd = 1.5, col = "maroon")
text(MWPcrit_60[1], lfm_pct[1],
     labels = "62% LFM", pos = 3, col = "maroon", cex = 0.9)

lines(MWPcrit_77, lfm_pct, lwd = 1.5, col = "goldenrod")
text(MWPcrit_77[1], lfm_pct[1],
     labels = "77% LFM", pos = 3, col = "goldenrod", cex = 0.9)

lines(MWPcrit_88, lfm_pct_green, lwd = 1.5, col = "forestgreen")
text(MWPcrit_88[1], lfm_pct_green[1],
     labels = "88% LFM", pos = 4, col = "forestgreen", cex = 0.9)

# ----------------------------
# Plot points (colored by growth form)
# ----------------------------
points(
  x = spres$minWP,
  y = (1 / spres$model1.int) * 100,
  pch = 16,
  cex = 1.1,
  col = color_growth[spres$growth_form] # Mapping to growth form colors
)

# ----------------------------
# Labels for selected species (one label per species)
# ----------------------------
label_spp <- c("PSEMEN", "UMBCAL", "ABCO", "PIJE", "PINJEF", "ABICON",
               "CALDEC")

special <- c("PIJE", "PINJEF", "CALDEC", "QUEKEL", "CEACOR")

spres_lab <- spres %>%
  filter(SpCode6 %in% label_spp) %>%
  mutate(
    x = minWP,
    y = (1 / model1.int) * 100
  ) %>%
  group_by(SpCode6) %>%
  slice(1) %>%
  ungroup()

# right-side labels
spres_lab_other <- spres_lab %>% filter(!SpCode6 %in% special)
if (nrow(spres_lab_other) > 0) {
  text(
    x = spres_lab_other$x,
    y = spres_lab_other$y,
    labels = spres_lab_other$SpCode6,
    pos = 4,
    offset = 0.3,
    cex = 0.75,
    col = species_cols[spres_lab_other$SpCode6]
  )
}

# special labels above point
spres_lab_special <- spres_lab %>% filter(SpCode6 %in% special)
if (nrow(spres_lab_special) > 0) {
  text(
    x = spres_lab_special$x,
    y = spres_lab_special$y,
    labels = spres_lab_special$SpCode6,
    pos = 3,
    offset = 0.45,
    adj = c(0.5, 0),
    cex = 0.75,
    col = species_cols[spres_lab_special$SpCode6]
  )
}

# ----------------------------
# Legend (to the right, updated for Growth Form)
# ----------------------------
usr <- par("usr")   # c(xmin, xmax, ymin, ymax)
pin <- par("pin")   # plot dimensions in inches
x_per_inch <- diff(usr[1:2]) / pin[1]
x_shift <- 0.9 * x_per_inch  # ~0.9 inch to the right

legend(
  x = usr[2] + x_shift,
  y = usr[4],
  xjust = 1, yjust = 1,
  legend = c("Tree", "Shrub"), # Capitalized, publication-ready labels
  pch = 16,
  col = color_growth[c("tree", "shrub")],
  bty = "n",
  cex = 0.8,
  title = "Growth Form"
)

par(op)
dev.off()
```





