---
title: "Untitled"
author: "Indra Boving"
date: "2026-01-27"
output: html_document
---

Use bayesian modeling to model iLFM ~ psi_md

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(brms)
library(tidyverse)
library(here)
```


2) If you were to take all the significant relationships of model parameters with traits, could you build a multivariate model that predicts LFMmin, and then compare it to observed? The value would be for modelers who have functional trait databases, to show how close we can get in predictions based only on the functional traits

#Read in data: 

```{r}
concept_df <- read_csv(here("data", "concept_df.csv")) %>% 
  filter(group == "species") %>% 
  select(mwp, ilfm, species, study) 

q3_df <- read_csv(here("data", "q3_df.csv"))

q4_df <- merge(concept_df, q3_df, by = c("species")) %>%
  filter(!is.na(sla)) %>%
  mutate(
    psi_z = as.numeric(scale(mwp)),   # or psi_md if that’s the column name
    sla_z = as.numeric(scale(sla))
  )

q4_df %>% write_csv(here("data", "q4_df.csv"))
```
#Table SX: Model comparison table

No interactions:
```{r, eval = F}
library(dplyr)
library(lme4)
library(lmerTest)
library(Metrics)
library(here)

# ---------------------------------------------------------
# 1. Load and create the "Master" dataset
# ---------------------------------------------------------
df <- q4_df

# Drop NAs for ALL variables used across ALL models
df_mod <- df %>%
  mutate(max_lfm = intercept) %>% 
  filter(plc50_stem >= -8 & plc50_stem <= -1) %>%
  filter(sla > 0 & sla < 100) %>% 
  filter(
    !is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), 
    !is.na(functional_type), !is.na(psi_tlp), !is.na(ldmc), 
    !is.na(max_lfm), !is.na(min_wp), !is.na(species)
  ) %>%
  mutate(
    mwp_z    = as.numeric(scale(mwp)),
    p50_z    = as.numeric(scale(plc50_stem)),
    sla_z    = as.numeric(scale(sla)),
    tlp_z    = as.numeric(scale(psi_tlp)),
    ldmc_z   = as.numeric(scale(ldmc)),
    lfmmax_z = as.numeric(scale(max_lfm)),
    psimin_z = as.numeric(scale(min_wp)),
    functional_type = as.factor(functional_type)
  )

# ---------------------------------------------------------
# 2. Automatically Generate EVERY Model Formula (128 Total)
# ---------------------------------------------------------
# The 7 variables we are testing:
candidate_vars <- c("p50_z", "sla_z", "ldmc_z", "tlp_z", "lfmmax_z", "psimin_z", "functional_type")

formulas <- list()

# Add the Base Model first
formulas[["Base (WP Only)"]] <- as.formula("ilfm ~ mwp_z + (1 | species)")

# Loop through and create every 1-trait, 2-trait... up to 7-trait combination
for (i in 1:length(candidate_vars)) {
  
  # Get all unique combinations of length i
  combos <- combn(candidate_vars, i, simplify = FALSE)
  
  for (combo in combos) {
    # 1. Create a clean name for the output table
    clean_name <- paste(combo, collapse = " + ")
    clean_name <- gsub("_z", "", clean_name) # Remove the _z for cleaner reading
    clean_name <- gsub("functional_type", "F.Type", clean_name)
    
    # 2. Build the formula string
    interaction_terms <- paste0("mwp_z * ", combo)
    formula_string <- paste("ilfm ~", paste(interaction_terms, collapse = " + "), "+ (1 | species)")
    
    # 3. Save to our formulas list
    formulas[[clean_name]] <- as.formula(formula_string)
  }
}

# ---------------------------------------------------------
# 3. Loop through and evaluate every model
# ---------------------------------------------------------
results_list <- list()
species_list <- unique(df_mod$species)

cat("Starting model evaluation. Testing 128 models... this may take a couple of minutes!\n")

for(model_name in names(formulas)) {
  
  current_formula <- formulas[[model_name]]
  
  # A. Fit the full model (Suppress warnings for over-parameterized combinations)
  full_mod <- suppressWarnings(lmer(current_formula, data = df_mod, REML = FALSE))
  
  # B. Get AIC
  mod_aic <- AIC(full_mod)
  
  # C. Get In-Sample R-squared
  in_sample_preds <- predict(full_mod)
  in_sample_r2 <- cor(df_mod$ilfm, in_sample_preds)^2
  
  # D. Run the Leave-One-Species-Out Cross Validation
  cv_preds <- numeric(nrow(df_mod))
  
  for(sp in species_list) {
    train_data <- df_mod %>% filter(species != sp)
    test_data  <- df_mod %>% filter(species == sp)
    
    cv_mod <- suppressWarnings(lmer(current_formula, data = train_data, REML = FALSE))
    cv_preds[df_mod$species == sp] <- predict(cv_mod, newdata = test_data, allow.new.levels = TRUE)
  }
  
  cv_r2 <- cor(df_mod$ilfm, cv_preds)^2
  
  # Save the results
  results_list[[model_name]] <- data.frame(
    Model = model_name,
    AIC = mod_aic,
    In_Sample_R2 = in_sample_r2,
    Cross_Val_R2 = cv_r2
  )
}

# ---------------------------------------------------------
# 4. Format and export the final table
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

# Calculate Delta AIC (Difference from the best model)
min_aic <- min(model_comparison_df$AIC)
model_comparison_df <- model_comparison_df %>%
  mutate(
    Delta_AIC = AIC - min_aic
  ) %>%
  # Arrange so the best models (lowest AIC) are at the top
  arrange(Delta_AIC) %>%
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  select(Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

cat("Finished! Exporting table.\n")

# Print the top 10 models to console
print(head(model_comparison_df, 10))

# Save to CSV
write.csv(model_comparison_df, here("figures", "Table_Model_Comparison_All_Subsets_notints.csv"), row.names = FALSE)
```

Interactions: 
```{r, eval = F}
library(dplyr)
library(lme4)
library(lmerTest)
library(Metrics)
library(here)
# install.packages("MuMIn") # Run this once if you don't have it!
library(MuMIn) 

# ---------------------------------------------------------
# 1. Load and create the "Master" dataset
# ---------------------------------------------------------
df <- q4_df

# Drop NAs for ALL variables used across ALL models
df_mod <- df %>%
  mutate(max_lfm = intercept) %>% 
  filter(plc50_stem >= -8 & plc50_stem <= -1) %>%
  filter(sla > 0 & sla < 100) %>% 
  filter(
    !is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), 
    !is.na(functional_type), !is.na(psi_tlp), !is.na(ldmc), 
    !is.na(max_lfm), !is.na(min_wp), !is.na(species)
  ) %>%
  mutate(
    mwp_z    = as.numeric(scale(mwp)),
    p50_z    = as.numeric(scale(plc50_stem)),
    sla_z    = as.numeric(scale(sla)),
    tlp_z    = as.numeric(scale(psi_tlp)),
    ldmc_z   = as.numeric(scale(ldmc)),
    lfmmax_z = as.numeric(scale(max_lfm)),
    psimin_z = as.numeric(scale(min_wp)),
    functional_type = as.factor(functional_type)
  )

# ---------------------------------------------------------
# 2. Define the Global Model and Run Dredge
# ---------------------------------------------------------
# na.fail is strictly required for MuMIn's dredge function to run
options(na.action = "na.fail") 

# This is the "Kitchen Sink" model. 
# Writing mwp_z * (trait1 + trait2...) tells the model to evaluate both
# the main effects AND their interactions with water potential.
global_mod <- suppressWarnings(lmer(
  ilfm ~ mwp_z * (p50_z + sla_z + ldmc_z + tlp_z + lfmmax_z + psimin_z + functional_type) + (1 | species),
  data = df_mod, 
  REML = FALSE
))

cat("Dredging 2,187 possible models to find the best combinations (this is fast)...\n")
# dredge() automatically generates and tests every statistically valid sub-model
model_set <- suppressWarnings(dredge(global_mod))

# ---------------------------------------------------------
# 3. Filter Top Models and Run Cross-Validation
# ---------------------------------------------------------
# We only want to run the heavy CV loop on the models that are actually good.
# subset = delta <= 10 grabs models within 10 AIC of the absolute best one.
top_models <- get.models(model_set, subset = delta <= 10)

# Make sure we also manually include the Base Model for comparison!
base_mod <- suppressWarnings(lmer(ilfm ~ mwp_z + (1 | species), data = df_mod, REML = FALSE))
top_models[["Base (WP Only)"]] <- base_mod

results_list <- list()
species_list <- unique(df_mod$species)

cat(sprintf("Running Leave-One-Species-Out CV on the top %d models. This will take a minute...\n", length(top_models)))

for(i in seq_along(top_models)) {
  
  current_mod <- top_models[[i]]
  current_formula <- formula(current_mod)
  
  # A. Get AIC
  mod_aic <- AIC(current_mod)
  
  # B. Get In-Sample R-squared
  in_sample_preds <- predict(current_mod)
  in_sample_r2 <- cor(df_mod$ilfm, in_sample_preds)^2
  
  # C. Run the Leave-One-Species-Out Cross Validation
  cv_preds <- numeric(nrow(df_mod))
  
  for(sp in species_list) {
    train_data <- df_mod %>% filter(species != sp)
    test_data  <- df_mod %>% filter(species == sp)
    
    cv_mod <- suppressWarnings(lmer(current_formula, data = train_data, REML = FALSE))
    cv_preds[df_mod$species == sp] <- predict(cv_mod, newdata = test_data, allow.new.levels = TRUE)
  }
  cv_r2 <- cor(df_mod$ilfm, cv_preds)^2
  
  # D. Create a clean name for the output table
  if(names(top_models)[i] == "Base (WP Only)") {
    clean_name <- "Base (WP Only)"
  } else {
    # Extract the formula as a string and clean it up for readability
    clean_name <- Reduce(paste, deparse(current_formula[[3]]))
    clean_name <- gsub(" \\+ \\(1 \\| species\\)", "", clean_name)
    clean_name <- gsub("_z", "", clean_name)
    clean_name <- gsub("functional_type", "F.Type", clean_name)
    clean_name <- gsub(":", " x ", clean_name) # Makes interactions look nice (e.g. mwp x p50)
  }
  
  # Save the results
  results_list[[i]] <- data.frame(
    Model = clean_name,
    AIC = mod_aic,
    In_Sample_R2 = in_sample_r2,
    Cross_Val_R2 = cv_r2
  )
}

# ---------------------------------------------------------
# 4. Format and export the final table
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

# Calculate Delta AIC
min_aic <- min(model_comparison_df$AIC)
model_comparison_df <- model_comparison_df %>%
  mutate(
    Delta_AIC = AIC - min_aic
  ) %>%
  arrange(Delta_AIC) %>%
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  select(Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

cat("Finished! Exporting table.\n")

# Print the top 10 models to console
print(head(model_comparison_df, 10))

# Save to CSV
write.csv(model_comparison_df, here("figures", "Table_Model_Comparison_Dredged.csv"), row.names = FALSE)

```

```{r, eval = F}
library(dplyr)
library(stringr)

# ---------------------------------------------------------
# 1. Top Overall Models + Base Model
# ---------------------------------------------------------
top_overall <- model_comparison_df %>%
  # Keep models within 2 Delta AIC, OR keep the base model
  filter(Delta_AIC <= 2 | Model == "Base (WP Only)") %>%
  arrange(Delta_AIC)

# ---------------------------------------------------------
# 2. Top Trait-Only Models + Base Model
# ---------------------------------------------------------
top_traits_only <- model_comparison_df %>%
  # Filter out LFMmax and Psimin, BUT protect the Base Model
  filter((!str_detect(Model, "lfmmax") & !str_detect(Model, "psimin")) | Model == "Base (WP Only)") %>%
  # Recalculate Delta AIC relative to the best model in this group
  mutate(Delta_AIC_Subset = round(AIC - min(AIC), 1)) %>%
  # Keep only the top models AND the base model
  filter(Delta_AIC_Subset <= 2 | Model == "Base (WP Only)") %>%
  arrange(Delta_AIC_Subset)

top_traits_only

# ---------------------------------------------------------
# 3. Top Trait-Only Models WITH NO Functional Type + Base Model
# ---------------------------------------------------------
top_traits_no_ftype <- model_comparison_df %>%
  # Filter out LFMmax, Psimin, and F.Type, BUT protect the Base Model
  filter((!str_detect(Model, "lfmmax") & 
          !str_detect(Model, "psimin") & 
          !str_detect(Model, "F.Type")) | Model == "Base (WP Only)") %>%
  # Recalculate Delta AIC relative to the best model in this group
  mutate(Delta_AIC_Subset = round(AIC - min(AIC), 1)) %>%
  # Keep only the top models AND the base model
  filter(Delta_AIC_Subset <= 2 | Model == "Base (WP Only)") %>%
  arrange(Delta_AIC_Subset)

# ---------------------------------------------------------
# View the Results
# ---------------------------------------------------------
cat("\n--- TOP OVERALL MODELS vs BASE ---\n")
print(top_overall)

cat("\n--- TOP TRAIT-ONLY MODELS vs BASE ---\n")
print(top_traits_only)

cat("\n--- TOP TRAIT-ONLY, NO F-TYPE MODELS vs BASE ---\n")
print(top_traits_no_ftype)

# Save them to distinct CSVs
write.csv(top_overall, here("figures", "Table_Top_Overall.csv"), row.names = FALSE)
write.csv(top_traits_only, here("figures", "Table_Top_Traits.csv"), row.names = FALSE)
write.csv(top_traits_no_ftype, here("figures", "Table_Top_Traits_No_FType.csv"), row.names = FALSE)
```

####Nice table:

```{r}
library(dplyr)
library(knitr)
library(kableExtra)
library(stringr)

# ---------------------------------------------------------
# 1. Prepare and Tag Each Subset 
# ---------------------------------------------------------
df_full <- top_overall %>%
  filter(Model != "Base (WP Only)") %>%
  mutate(Model_Group = "Full Models") %>%
  select(Model_Group, Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

df_traits_ftype <- top_traits_only %>%
  filter(Model != "Base (WP Only)") %>%
  mutate(Model_Group = "Trait Only + F. Type") %>%
  select(Model_Group, Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

df_traits <- top_traits_no_ftype %>%
  filter(Model != "Base (WP Only)") %>%
  mutate(Model_Group = "Trait Only") %>%
  select(Model_Group, Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

df_base <- model_comparison_df %>%
  filter(Model == "Base (WP Only)") %>%
  mutate(Model_Group = "Baseline") %>%
  select(Model_Group, Model, AIC, Delta_AIC, In_Sample_R2, Cross_Val_R2)

# ---------------------------------------------------------
# 2. Combine, Clean Names, and Sort
# ---------------------------------------------------------
combined_df <- bind_rows(df_full, df_traits_ftype, df_traits, df_base) %>%
  mutate(Model_Group = factor(Model_Group, levels = c("Full Models", "Trait Only + F. Type", "Trait Only", "Baseline"))) %>%
  arrange(Model_Group, AIC) %>%
  
  # Clean up the trait and parameter names to look professional
  mutate(
    Model = str_replace_all(Model, "p50", "P50"),
    Model = str_replace_all(Model, "sla", "SLA"),
    Model = str_replace_all(Model, "tlp", "TLP"),
    Model = str_replace_all(Model, "ldmc", "LDMC"),
    Model = str_replace_all(Model, "lfmmax", "Max LFM"),
    Model = str_replace_all(Model, "psimin", "Min WP"),
    Model = str_replace_all(Model, "F.Type", "Functional Type"),
    Model = str_replace_all(Model, "Base \\(WP Only\\)", "WP Only")
  ) %>%
  
  # Clean up column headers
  rename(
    `Model Group` = Model_Group,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

# Explicitly strip out any lingering row names to prevent the phantom column
rownames(combined_df) <- NULL

# ---------------------------------------------------------
# 3. Create the Publication-Ready kable Table
# ---------------------------------------------------------
# Added row.names = FALSE to ensure the row numbers don't print
pub_table <- kable(combined_df, format = "html", row.names = FALSE, escape = FALSE, align = "llcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  collapse_rows(columns = 1, valign = "top") 

# Print the table to the Viewer Pane
pub_table

library(kableExtra)
library(here)

# Save the formatted table as an HTML file
save_kable(pub_table, file = here("figures", "Table_Model_Comparison_Formatted_withTLP.html"))
```

####Check VIF: 
```{r}
library(dplyr)
library(car) # For checking VIF

# 1. Simple Correlation Check
# We want species-level correlation, so we drop duplicates first
species_traits <- df_mod %>%
  select(species, sla, ldmc) %>%
  distinct()

# Calculate Pearson correlation
cor(species_traits$sla, species_traits$ldmc) 
# You should get roughly -0.55

# 2. VIF Check on the Model
# Let's look at the model that has both SLA and LDMC
mod_vif_check <- lmer(
  ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species), 
  data = df_mod, 
  REML = TRUE
)

# Check the Variance Inflation Factor
vif(mod_vif_check)
```

#Table S5: Final model selection
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(Metrics)
library(knitr)
library(kableExtra)
library(here)

# ---------------------------------------------------------
# 1. Prepare the Master Dataset
# ---------------------------------------------------------
# We must drop NAs for all variables used in ANY of the 9 models
# so they are all compared on the exact same dataset.
df_mod <- q4_df %>%
  filter(plc50_stem >= -8 & plc50_stem <= -1) %>%
  filter(sla > 0 & sla < 100) %>% 
  filter(
    !is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), 
    !is.na(ldmc), !is.na(psi_tlp), !is.na(functional_type), !is.na(species)
  ) %>%
  mutate(
    mwp_z    = as.numeric(scale(mwp)),
    p50_z    = as.numeric(scale(plc50_stem)),
    sla_z    = as.numeric(scale(sla)),
    tlp_z    = as.numeric(scale(psi_tlp)),
    ldmc_z   = as.numeric(scale(ldmc)),
    functional_type = as.factor(functional_type)
  )

# ---------------------------------------------------------
# 2. Define Your 9 A Priori Models
# ---------------------------------------------------------
formulas <- list(
  "P50 + SLA + LDMC + TLP" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * tlp_z + (1 | species),
  
  "P50 + SLA + LDMC + TLP + F. Type" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * tlp_z + (1 | species),
  
  "P50 + SLA + LDMC + F. Type" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "TLP + SLA + LDMC + F. Type" = ilfm ~ mwp_z * tlp_z + mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "P50 + SLA + LDMC" = ilfm ~ mwp_z * p50_z + mwp_z * sla_z + mwp_z * ldmc_z + (1 | species),
  
  "SLA + LDMC + F. Type" = ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + (1 | species),
  
  "LDMC + F. Type" = ilfm ~ mwp_z * ldmc_z +  functional_type * mwp_z + (1 | species),
  
  "SLA + F. Type" = ilfm ~ mwp_z * sla_z + functional_type * mwp_z + (1 | species),
  
  "SLA Only" = ilfm ~ mwp_z * sla_z + (1 | species),
  
  # P50 + TLP + F. Type (Hydraulics only, no leaf economics)
  "TLP + P50 + F. Type" = lmer(ilfm ~ mwp_z * p50_z + mwp_z * tlp_z + functional_type * mwp_z + (1 | species), df_mod),
  
  "TLP Only" = ilfm ~ mwp_z * tlp_z + (1 | species),
  
  "P50 Only" = ilfm ~ mwp_z * p50_z+ (1 | species),
  
  "LDMC Only" = ilfm ~ mwp_z * ldmc_z+ (1 | species),
  
  "F.Type Only" = ilfm ~ mwp_z * functional_type + (1 | species),
  
  "WP Only (Null)" = ilfm ~ mwp_z + (1 | species)
)

# ---------------------------------------------------------
# 3. Evaluate Models and Run Cross-Validation
# ---------------------------------------------------------
results_list <- list()
species_list <- unique(df_mod$species)

cat("Running evaluations and cross-validation for 9 hypothesis-driven models...\n")

for(model_name in names(formulas)) {
  
  current_formula <- formulas[[model_name]]
  
  # A. Fit Full Model (REML = FALSE is required to compare fixed effects with AIC)
  full_mod <- lmer(current_formula, data = df_mod, REML = FALSE)
  
  # B. Calculate AIC & In-Sample R2
  mod_aic <- AIC(full_mod)
  in_sample_preds <- predict(full_mod)
  in_sample_r2 <- cor(df_mod$ilfm, in_sample_preds)^2
  
  # C. Leave-One-Species-Out CV Loop
  cv_preds <- numeric(nrow(df_mod))
  
  for(sp in species_list) {
    train_data <- df_mod %>% filter(species != sp)
    test_data  <- df_mod %>% filter(species == sp)
    
    cv_mod <- lmer(current_formula, data = train_data, REML = FALSE)
    cv_preds[df_mod$species == sp] <- predict(cv_mod, newdata = test_data, allow.new.levels = TRUE)
  }
  cv_r2 <- cor(df_mod$ilfm, cv_preds)^2
  
  # D. Store Results
  results_list[[model_name]] <- data.frame(
    Model = model_name,
    AIC = mod_aic,
    In_Sample_R2 = in_sample_r2,
    Cross_Val_R2 = cv_r2
  )
}

# ---------------------------------------------------------
# 4. Format the Dataframe
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

min_aic <- min(model_comparison_df$AIC)

final_df <- model_comparison_df %>%
  mutate(Delta_AIC = AIC - min_aic) %>%
  arrange(Delta_AIC) %>%
  # Round numbers for clean presentation
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  # Clean up headers
  rename(
    `Candidate Model` = Model,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

rownames(final_df) <- NULL

# ---------------------------------------------------------
# 5. Create and Save the Publication kable Table
# ---------------------------------------------------------
pub_table <- kable(final_df, format = "html", escape = FALSE, align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  # Highlight the best model row in bold
  row_spec(1, bold = TRUE) 

# Print to Viewer
pub_table

# Save as an HTML file for easy copy/pasting into Word
save_kable(pub_table, file = here("figures", "Table_Hypothesis_Models.html"))

cat("Done! Table saved to figures/Table_Hypothesis_Models.html\n")

# ---------------------------------------------------------
# 4. Filter, Format, and Rank
# ---------------------------------------------------------
model_comparison_df <- do.call(rbind, results_list)

min_aic <- min(model_comparison_df$AIC)

final_df <- model_comparison_df %>%
  mutate(Delta_AIC = AIC - min_aic) %>%
  # Explicitly keep models within 10 AIC, PLUS your two baselines
  filter(Delta_AIC <= 10 | Model %in% c("WP Only (Null)", "F. Type Only")) %>%
  arrange(Delta_AIC) %>%
  # Round numbers for clean presentation
  mutate(
    AIC = round(AIC, 1),
    Delta_AIC = round(Delta_AIC, 1),
    In_Sample_R2 = round(In_Sample_R2, 3),
    Cross_Val_R2 = round(Cross_Val_R2, 3)
  ) %>%
  # Clean up headers
  rename(
    `Candidate Model` = Model,
    `Delta AIC` = Delta_AIC,
    `In-Sample R-squared` = In_Sample_R2,
    `Cross-Val R-squared` = Cross_Val_R2
  )

rownames(final_df) <- NULL

# ---------------------------------------------------------
# 5. Create and Save the Publication kable Table
# ---------------------------------------------------------
pub_table <- kable(final_df, format = "html", escape = FALSE, align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "left") %>%
  row_spec(1, bold = TRUE) 

# Print to Viewer
pub_table

# Save as an HTML file for easy copy/pasting into Word
save_kable(pub_table, file = here("figures", "Table_Hypothesis_Models_Filtered.html"))

cat("Done! Table saved to figures/Table_Hypothesis_Models_Filtered.html\n")
```

#Fig. 10
####v1: With just P50, LDMC, SLA: 

```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(Metrics) # for calculating RMSE and R-squared

# ---------------------------------------------------------
# 1. Load and clean data
# ---------------------------------------------------------
df <- q4_df

df_mod <- df %>%
  # Drop missing values, now including ldmc
  filter(!is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), !is.na(ldmc), !is.na(species)) %>%
  # Standardize continuous predictors (z-scores)
  mutate(
    mwp_z = as.numeric(scale(mwp)),
    p50_z = as.numeric(scale(plc50_stem)),
    sla_z = as.numeric(scale(sla)),
    ldmc_z = as.numeric(scale(ldmc))
  )

# ---------------------------------------------------------
# 2. Frequentist Framework (lmer)
# ---------------------------------------------------------
# Model with interactions for P50, SLA, and LDMC
freq_mod <- lmer(
  ilfm ~  mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * p50_z + (1 | species), 
  data = df_mod, 
  REML = TRUE
)

# summary() will now give you p-values
summary(freq_mod)

# ---------------------------------------------------------
# 3. In-Sample Observed vs. Predicted (The full model)
# ---------------------------------------------------------
df_mod$predicted_ilfm <- predict(freq_mod)

in_sample_r2 <- cor(df_mod$ilfm, df_mod$predicted_ilfm)^2

plot_insample <- ggplot(df_mod, aes(x = predicted_ilfm, y = ilfm)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", linewidth = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Model Fit (All Data)",
    subtitle = paste0("R-squared: ", round(in_sample_r2, 2)),
    x = "Predicted iLFM",
    y = "Observed iLFM"
  ) +
  theme_classic(base_size = 14)

# ---------------------------------------------------------
# 4. Leave-One-Species-Out Cross Validation
# ---------------------------------------------------------
df_mod$cv_predicted_ilfm <- NA
species_list <- unique(df_mod$species)

for(sp in species_list) {
  
  train_data <- df_mod %>% filter(species != sp)
  test_data <- df_mod %>% filter(species == sp)
  
  temp_mod <- lmer(
    ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + mwp_z * p50_z + (1 | species), 
    data = train_data,
    REML = FALSE 
  )
  
  preds <- predict(temp_mod, newdata = test_data, allow.new.levels = TRUE)
  df_mod$cv_predicted_ilfm[df_mod$species == sp] <- preds
}

cv_r2 <- cor(df_mod$ilfm, df_mod$cv_predicted_ilfm)^2

plot_cv <- ggplot(df_mod, aes(x = cv_predicted_ilfm, y = ilfm)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", linewidth = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Cross-Validation (Leave-One-Species-Out)",
    subtitle = paste0("Predictive R-squared: ", round(cv_r2, 2)),
    x = "Predicted iLFM (Unseen Species)",
    y = "Observed iLFM"
  ) +
  theme_classic(base_size = 14)

# ---------------------------------------------------------
# 5. View them side-by-side
# ---------------------------------------------------------
plot_insample + plot_cv


library(ggplot2)
library(patchwork)
library(Metrics)
library(here)

# ---------------------------------------------------------
# 1. Calculate Metrics
# ---------------------------------------------------------
# Assuming in_sample_r2 and cv_r2 are already calculated from earlier steps
in_sample_rmse <- rmse(df_mod$ilfm, df_mod$predicted_ilfm)
cv_rmse <- rmse(df_mod$ilfm, df_mod$cv_predicted_ilfm)

# Find the absolute min and max across ALL predictions and observations 
# so both panels have the exact same axis scale
min_val <- min(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)
max_val <- max(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)

# Create the annotation text for both
anno_text_insample <- sprintf("In-Sample R2 = %.2f\nRMSE = %.2f", in_sample_r2, in_sample_rmse)
anno_text_cv <- sprintf("Predictive R2 = %.2f\nRMSE = %.2f", cv_r2, cv_rmse)

# ---------------------------------------------------------
# 2. Build In-Sample Plot (Panel A)
# ---------------------------------------------------------
plot_insample_pub <- ggplot(df_mod, aes(x = predicted_ilfm, y = ilfm)) +
  geom_point(alpha = 0.5, shape = 21, fill = "grey50", color = "black", size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_insample, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (In-Sample)",
    y = "Observed iLFM"
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) 
  )

# ---------------------------------------------------------
# 3. Build Cross-Validation Plot (Panel B)
# ---------------------------------------------------------
plot_cv_pub <- ggplot(df_mod, aes(x = cv_predicted_ilfm, y = ilfm)) +
  geom_point(alpha = 0.5, shape = 21, fill = "grey50", color = "black", size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_cv, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (Unseen Species)",
    y = "Observed iLFM"
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) 
  )

# ---------------------------------------------------------
# 4. Combine and Save
# ---------------------------------------------------------
# Stitch together with patchwork and add A/B tags
combined_plot <- plot_insample_pub + plot_cv_pub + 
  plot_annotation(tag_levels = 'A')

combined_plot

ggsave(
  filename = here("figures", "Fig10.png"), 
  plot     = combined_plot,
  width    = 12,   # Doubled the width to fit both squares
  height   = 6,   
  units    = "in",
  dpi      = 300
)
```

####v2: Colored points 
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(patchwork)
library(Metrics) # for calculating RMSE and R-squared
library(here)

# ---------------------------------------------------------
# 1. Load, clean data, and assign plot groups
# ---------------------------------------------------------
df <- q4_df

df_mod <- df %>%
  # Drop missing values
  filter(!is.na(mwp), !is.na(ilfm), !is.na(plc50_stem), !is.na(sla), !is.na(ldmc), !is.na(species), !is.na(psi_tlp)) %>%
  # Standardize continuous predictors (z-scores)
  mutate(
    mwp_z = as.numeric(scale(mwp)),
    p50_z = as.numeric(scale(plc50_stem)),
    sla_z = as.numeric(scale(sla)),
    tlp_z = as.numeric(scale(psi_tlp)),
    ldmc_z = as.numeric(scale(ldmc))
  ) %>%
  # Create the grouping variable for the plot colors
  mutate(
    plot_group = case_when(
      grepl("gymnosperm", functional_type, ignore.case = TRUE) ~ "Conifer",
      grepl("sal", species, ignore.case = TRUE) ~ "Salvia",
      TRUE ~ "Other Angiosperm"
    ),
    # Make it a factor to control the legend order
    plot_group = factor(plot_group, levels = c("Conifer", "Salvia", "Other Angiosperm"))
  )

# ---------------------------------------------------------
# 2. Frequentist Framework (lmer)
# ---------------------------------------------------------
freq_mod <- lmer(
  ilfm ~  mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * p50_z + mwp_z * tlp_z+ (1 | species), 
  data = df_mod, 
  REML = TRUE
)

# summary() will now give you p-values
summary(freq_mod)

# ---------------------------------------------------------
# 3. Model Predictions (In-Sample and Out-of-Sample)
# ---------------------------------------------------------
# In-Sample
df_mod$predicted_ilfm <- predict(freq_mod)
in_sample_r2 <- cor(df_mod$ilfm, df_mod$predicted_ilfm)^2
in_sample_rmse <- rmse(df_mod$ilfm, df_mod$predicted_ilfm)

# Out-of-Sample (Leave-One-Species-Out)
df_mod$cv_predicted_ilfm <- NA
species_list <- unique(df_mod$species)

for(sp in species_list) {
  train_data <- df_mod %>% filter(species != sp)
  test_data <- df_mod %>% filter(species == sp)
  
  temp_mod <- lmer(
    ilfm ~ mwp_z * sla_z + mwp_z * ldmc_z + functional_type * mwp_z + mwp_z * p50_z + mwp_z * tlp_z +(1 | species), 
    data = train_data,
    REML = FALSE 
  )
  
  preds <- predict(temp_mod, newdata = test_data, allow.new.levels = TRUE)
  df_mod$cv_predicted_ilfm[df_mod$species == sp] <- preds
}

cv_r2 <- cor(df_mod$ilfm, df_mod$cv_predicted_ilfm)^2
cv_rmse <- rmse(df_mod$ilfm, df_mod$cv_predicted_ilfm)

# ---------------------------------------------------------
# 4. Build Publication-Ready Plots
# ---------------------------------------------------------
# Find the absolute min and max across ALL predictions and observations 
min_val <- min(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)
max_val <- max(c(df_mod$ilfm, df_mod$predicted_ilfm, df_mod$cv_predicted_ilfm), na.rm = TRUE)

anno_text_insample <- sprintf("In-Sample R2 = %.2f\nRMSE = %.2f", in_sample_r2, in_sample_rmse)
anno_text_cv <- sprintf("Predictive R2 = %.2f\nRMSE = %.2f", cv_r2, cv_rmse)

# Custom color palette
my_colors <- c(
  "Conifer" = "#636B2F",       # Light Green
  "Salvia" = "#F768A1",           # Pink
  "Other Angiosperm" = "grey70"   # Default Grey
)
 df_mod <- df_mod %>%
  # Make sure the label is "Conifer" instead of "Gymnosperm"
  mutate(
    plot_group = str_replace(plot_group, "Gymnosperm", "Conifer"),
    plot_group = factor(plot_group, levels = c("Conifer", "Salvia", "Other Angiosperm"))
  ) %>%
  # Sorting by desc(plot_group) puts "Other Angiosperm" first in the dataframe
  arrange(desc(plot_group))
# Panel A: In-Sample Plot
plot_insample_pub <- ggplot(df_mod, aes(x = predicted_ilfm, y = ilfm)) +
  # Map the fill to our new plot_group
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_insample, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (In-Sample)",
    y = "Observed iLFM",
    fill = "Plant Group" # Clean legend title
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = "none" # Hide legend on the first plot so it doesn't duplicate
  )

# Panel B: Cross-Validation Plot
plot_cv_pub <- ggplot(df_mod, aes(x = cv_predicted_ilfm, y = ilfm)) +
  geom_point(aes(fill = plot_group), alpha = 0.6, shape = 21, color = "black", size = 2.5) +
  scale_fill_manual(values = my_colors) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed", linewidth = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "#2C7FB8", linewidth = 1.2) +
  annotate("text", x = min_val, y = max_val, label = anno_text_cv, 
           hjust = 0, vjust = 1, size = 5) +
  labs(
    x = "Predicted iLFM (Unseen Species)",
    y = "Observed iLFM",
    fill = "Plant Group"
  ) +
  coord_fixed(xlim = c(min_val, max_val), ylim = c(min_val, max_val)) +
  theme_classic(base_size = 14) +
  theme(
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) 
  )

# ---------------------------------------------------------
# 5. Combine and Save
# ---------------------------------------------------------
# Stitch together with patchwork and add A/B tags
# collect_guides() ensures we only have one clean legend for the whole figure
combined_plot <- plot_insample_pub + plot_cv_pub + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = "collect") 

combined_plot

ggsave(
  filename = here("figures", "Fig10_ftype.png"), 
  plot     = combined_plot,
  width    = 13,   # Added an extra inch of width to make room for the legend
  height   = 6,   
  units    = "in",
  dpi      = 300
)
```

##Bayesian approach: 

```{r, eval = F}
#remove.packages("rstan")
#if (file.exists(".RData")) file.remove(".RData")

Sys.setenv(DOWNLOAD_STATIC_LIBV8 = 1) # only necessary for Linux without the nodejs library / headers
install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)
example(stan_model, package = "rstan", run.dontrun = TRUE)

library("rstan") # observe startup messages
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

install.packages(c("coda","mvtnorm","devtools","loo","dagitty"))
install.packages("pkgbuild")
devtools::install_github("rmcelreath/rethinking")

library(rethinking)
```


```{r, eval = F}
# ---------------------------------------------------------
# 3. Bayesian Framework (ulam / rethinking)
# ---------------------------------------------------------
# Prepare a clean list for the ulam function
dat_list <- list(
  ilfm = df_mod$ilfm,
  mwp = df_mod$mwp_z,
  p50 = df_mod$p50_z,
  sla = df_mod$sla_z,
  # rethinking requires group variables to be 1-indexed integers
  sp_id = as.integer(as.factor(df_mod$species)) 
)

# Build the Bayesian model
bayes_mod <- ulam(
  alist(
    # Likelihood
    ilfm ~ dnorm(mu, sigma),
    
    # Linear model: intercepts vary by species. 
    # Slopes vary based on P50 and SLA.
    mu <- a[sp_id] + 
          (b_mwp + b_mwp_p50 * p50 + b_mwp_sla * sla) * mwp + 
          b_p50 * p50 + 
          b_sla * sla,
    
    # Adaptive prior for species random intercepts
    a[sp_id] ~ dnorm(a_bar, sigma_sp),
    a_bar ~ dnorm(0.5, 0.5), # Assuming iLFM centers roughly around 0.5; adjust if needed
    sigma_sp ~ dexp(1),
    
    # Fixed effect priors (Weakly informative because data is z-scored)
    b_mwp ~ dnorm(0, 1),
    b_p50 ~ dnorm(0, 1), 
    b_sla ~ dnorm(0, 1),
    
    # Interaction priors
    b_mwp_p50 ~ dnorm(0, 0.5),
    b_mwp_sla ~ dnorm(0, 0.5),
    
    # Residual variance
    sigma ~ dexp(1)
  ),
  data = dat_list,
  chains = 4, 
  cores = 4, 
  iter = 2000
)

# View the posterior estimates
precis(bayes_mod)
```
#Visualize bayesian approach: 

```{r, eval = F}
library(rethinking)
library(ggplot2)

# 1. Create a synthetic dataset to predict across
# We vary Water Potential from -2.5 to 2.5 SD
# We hold P50 at its mean (0)
# We test three levels of SLA: Low (-1), Mean (0), High (+1)
pred_dat <- expand.grid(
  mwp = seq(-2.5, 2.5, length.out = 30),
  p50 = 0,
  sla = c(-1, 0, 1),
  sp_id = 1 # Placeholder species ID
)

# 2. Generate posterior predictions
# replace = list(sp_id = a_bar) tells the model to ignore the placeholder species
# and just use the "grand average" intercept across all species
post_link <- link(bayes_mod, data = pred_dat, replace = list(sp_id = a_bar))

# 3. Extract the mean prediction and 89% Credible Intervals
pred_dat$predicted_ilfm <- apply(post_link, 2, mean)
pred_dat$ci_lower <- apply(post_link, 2, PI, prob = 0.89)[1, ]
pred_dat$ci_upper <- apply(post_link, 2, PI, prob = 0.89)[2, ]

# 4. Clean up labels for the plot
pred_dat$SLA_Level <- factor(
  pred_dat$sla, 
  levels = c(-1, 0, 1), 
  labels = c("Low SLA (-1 SD)", "Average SLA", "High SLA (+1 SD)")
)

# 5. Build the plot
ggplot(pred_dat, aes(x = mwp, y = predicted_ilfm, color = SLA_Level, fill = SLA_Level)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2, color = NA) +
  scale_color_manual(values = c("#2C7FB8", "grey50", "#D1495B")) +
  scale_fill_manual(values = c("#2C7FB8", "grey50", "#D1495B")) +
  labs(
    title = "Bayesian Posterior: Effect of SLA on Drydown Slope",
    x = "Midday Water Potential (z-score)",
    y = "Predicted iLFM",
    color = "SLA", fill = "SLA"
  ) +
  theme_classic(base_size = 14) +
  theme(legend.position = "bottom")
```
```{r, eval = F}
# Generate predictions for the actual data
in_sample_post <- link(bayes_mod)
df_mod$bayes_pred <- apply(in_sample_post, 2, mean)

# Calculate R-squared equivalent
bayes_r2 <- cor(df_mod$ilfm, df_mod$bayes_pred)^2

# Plot
ggplot(df_mod, aes(x = bayes_pred, y = ilfm)) +
  geom_point(alpha = 0.4, color = "grey30") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", linewidth = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Bayesian Model Fit (In-Sample)",
    subtitle = paste0("Bayesian R-squared: ", round(bayes_r2, 2)),
    x = "Predicted iLFM",
    y = "Observed iLFM"
  ) +
  theme_classic(base_size = 14)
```

